---
title: "Detergent soluble / insoluble proteomics (HK3-124)"
output: html_document
---

Publication figures for the soluble/insoluble unenriched proteomics run. Reads the CSVs
written by `solubility_analysis.ipynb`; does no computation of its own.

```{r setup, message=FALSE, warning=FALSE}
if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}
source("../../bin/figure_utils.R")

data_dir     <- "../../data/solubility/"
results_dir  <- paste0(data_dir, "03_results/")
volcano_dir  <- paste0(results_dir, "volcano_plots/")
gsea_dir     <- paste0(results_dir, "gsea/")

dir.create(paste0(volcano_dir, "figures"), recursive = TRUE, showWarnings = FALSE)
```

## Aesthetics

The ATP add-back palette (`notebooks/03_reactivity/reactivity_visualization.Rmd`) remapped onto
the six state x fraction groups: soluble keeps the base condition colour and a circle, insoluble
takes the accent colour and a diamond.

```{r}
solubility_fills <- c(
  "D2-Soluble"    = "darkgrey",
  "D2-Insoluble"  = "grey99",
  "D8A-Soluble"   = "#603EA6",
  "D8A-Insoluble" = "#D13B95",
  "D8C-Soluble"   = "#E27753",
  "D8C-Insoluble" = "#EF3D3B"
)

solubility_shapes <- c(
  "D2-Soluble"    = 21, "D2-Insoluble"  = 23,
  "D8A-Soluble"   = 21, "D8A-Insoluble" = 23,
  "D8C-Soluble"   = 21, "D8C-Insoluble" = 23
)

group_levels <- names(solubility_fills)

regulation_fills <- c(
  "Significant Up"          = SIG_UP_COL,
  "Significant Down"        = SIG_DOWN_COL,
  "Not Significant"         = UNCHANGED_COL,
  "Not Significant Up"      = UNCHANGED_COL,
  "Not Significant Down"    = UNCHANGED_COL,
  "Significant but <1.5 FC" = UNCHANGED_COL
)
```

`ggsave` silently drops points that fall outside a scale, so every limit in this document is
applied with `coord_cartesian()` and never with `xlim`/`ylim`/`scale_*_continuous(limits=)`.
This helper turns any "removed rows" warning into an error.

```{r}
save_checked <- function(plot, n_expected, fn, width, height, layers = 1) {
  built <- ggplot_build(plot)$data
  n_drawn <- sum(vapply(
    layers,
    function(i) sum(!is.na(built[[i]]$x) & !is.na(built[[i]]$y)),
    numeric(1)
  ))
  stopifnot(n_drawn == n_expected)

  withCallingHandlers(
    print(plot),
    warning = function(w) {
      if (grepl("Removed|missing values|non-finite", conditionMessage(w))) stop(w)
      invokeRestart("muffleWarning")
    }
  )
  save_plot(fn, width = width, height = height)
  invisible(plot)
}
```

## PCA

`replot_pca()` in `figure_utils.R` cannot be reused here: it hardcodes the five-colour
`cols` palette and references an undefined `results_dir` at two of its save points. This is a
local equivalent that keeps its geometry but adds the fraction shape scale and a legend.

```{r}
plot_solubility_pca <- function(pca_dir, out_name, x = "PC1", y = "PC2") {
  percent <- read_csv(paste0(pca_dir, "percent_explained.csv"), show_col_types = FALSE)
  pcs <- split(percent$percent_explained, percent$principal_component)

  scores <- read_csv(paste0(pca_dir, "pca_results.csv"), show_col_types = FALSE) %>%
    mutate(condition = factor(condition, levels = group_levels))

  pad <- function(v) {
    r <- range(v)
    r + c(-1, 1) * diff(r) * 0.12
  }

  plt <- scores %>%
    ggplot(aes(x = .data[[x]], y = .data[[y]],
               fill = condition, shape = condition)) +
    geom_point(size = POINT_SIZE, alpha = 0.9, col = "black", stroke = POINT_STROKE) +
    scale_fill_manual(values = solubility_fills, drop = FALSE) +
    scale_shape_manual(values = solubility_shapes, drop = FALSE) +
    coord_cartesian(xlim = pad(scores[[x]]), ylim = pad(scores[[y]])) +
    labs(x = paste0(x, " (", pcs[[x]], "%)"),
         y = paste0(y, " (", pcs[[y]], "%)")) +
    my_theme() +
    theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.position = "right",
      legend.key = element_blank(),
      legend.key.size = unit(0.25, "cm"),
      legend.margin = margin(l = -4),
      # my_theme() pulls legend text left by 5pt, which slides the label under the
      # key for names this long.
      legend.text = element_text(size = FONT_SIZE, colour = "black",
                                 family = FONT_FAMILY, margin = margin(l = 1))
    ) +
    guides(fill = guide_legend(override.aes = list(size = 1.6)))

  save_checked(plt, nrow(scores), paste0(pca_dir, out_name), width = 3.1, height = 2)
}
```

```{r fig.width=3.1, fig.height=2}
plot_solubility_pca(paste0(results_dir, "pca_unnormalized/"), "pca_plot")
```

The per-fraction constant is a translation on the log2 scale, so it loads onto PC1 and inflates
it (86.9% -> 94.8%) without moving the PC2 scores that separate the cell states. The
un-normalized PCA above is the one to use; the normalized version is kept for comparison.

```{r fig.width=3.1, fig.height=2}
plot_solubility_pca(paste0(results_dir, "pca_normalized/"), "pca_plot")
```

## Volcano plots

```{r}
volcano_data <- read_csv(paste0(volcano_dir, "long_volcano_df.csv"), show_col_types = FALSE)
protein_counts <- read_csv(paste0(volcano_dir, "volcano_protein_counts.csv"), show_col_types = FALSE)

# Shared axis ranges per family, so panels are comparable within a family without the
# off-centre fraction contrasts stretching the state contrasts.
family_limits <- volcano_data %>%
  group_by(family) %>%
  summarise(
    x_lo = min(log2_FC), x_hi = max(log2_FC),
    y_lo = 0,            y_hi = max(neg_log10_pval),
    .groups = "drop"
  ) %>%
  mutate(
    x_pad = (x_hi - x_lo) * 0.05,
    y_pad = y_hi * 0.08,
    x_lo = x_lo - x_pad, x_hi = x_hi + x_pad,
    y_hi = y_hi + y_pad
  )

family_limits
```

```{r}
# `data`, `counts`, `limits` and `out_dir` default to the HK3-124 globals defined above, so
# the original call sites are unchanged; the revision section passes its own.
plot_volcano <- function(condition_name, control_name,
                         data = volcano_data, counts = protein_counts,
                         limits = family_limits, out_dir = volcano_dir,
                         label_proteins = character()) {
  df <- data %>%
    filter(condition == condition_name, control_condition == control_name) %>%
    mutate(
      Regulation = factor(Regulation, levels = names(regulation_fills)),
      significant = Regulation %in% c("Significant Up", "Significant Down")
    )

  lims <- limits %>% filter(family == df$family[1])

  expected <- counts %>%
    filter(condition == condition_name, control_condition == control_name) %>%
    pull(n_tested)
  stopifnot(nrow(df) == expected)

  n_up   <- sum(df$Regulation == "Significant Up")
  n_down <- sum(df$Regulation == "Significant Down")

  plt <- ggplot(df, aes(x = log2_FC, y = neg_log10_pval,
                        fill = Regulation, alpha = significant)) +
    # Unchanged cloud underneath, significant points on top.
    geom_point(data = df %>% filter(!significant),
               size = 0.5, shape = 21, stroke = POINT_STROKE / 2, col = "black") +
    geom_point(data = df %>% filter(significant),
               size = 0.5, shape = 21, stroke = POINT_STROKE / 2, col = "black") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", linewidth = LINE_WIDTH) +
    geom_vline(xintercept = c(-FC_CUTOFF, FC_CUTOFF), linetype = "dashed",
               linewidth = LINE_WIDTH) +
    annotate("text", x = lims$x_lo, y = lims$y_hi, label = n_down,
             hjust = 0, vjust = 1, size = FONT_SIZE_MM, colour = SIG_DOWN_COL) +
    annotate("text", x = lims$x_hi, y = lims$y_hi, label = n_up,
             hjust = 1, vjust = 1, size = FONT_SIZE_MM, colour = SIG_UP_COL) +
    scale_fill_manual(values = regulation_fills, drop = FALSE, guide = "none") +
    scale_alpha_manual(values = c("FALSE" = 0.3, "TRUE" = 0.9), guide = "none") +
    coord_cartesian(xlim = c(lims$x_lo, lims$x_hi), ylim = c(lims$y_lo, lims$y_hi)) +
    labs(
      x = substitute(paste("protein expression, log"[2], "(", n, "/", m, ")"),
                     list(n = condition_name, m = control_name)),
      y = expression(paste("-log"[10], "(p-value)"))
    ) +
    my_theme() +
    theme(aspect.ratio = 1)

  # Proteins of interest are redrawn on top (already counted in layers 1-2, so the drawn-point
  # audit still sums to nrow(df)) with alpha forced to 1 so a non-significant one isn't faded.
  if (length(label_proteins)) {
    poi <- df %>% filter(protein %in% label_proteins)
    stopifnot(nrow(poi) == length(intersect(label_proteins, df$protein)))
    plt <- plt +
      geom_point(data = poi, size = 0.5, shape = 21, stroke = POINT_STROKE / 2,
                 colour = "black", alpha = 1) +
      geom_text_repel(
        data = poi, aes(x = log2_FC, y = neg_log10_pval, label = protein),
        inherit.aes = FALSE, size = FONT_SIZE_MM, colour = "black",
        box.padding = 0.35, min.segment.length = 0, segment.size = LINE_WIDTH)
  }

  # Wider than the repo's 2 in volcano default: the axis label carries two
  # <state>-<fraction> names, which is roughly twice the width of the D2/D8C names the
  # 2 in canvas was sized for, and it clips at that width.
  fn <- paste0(out_dir, "figures/", condition_name, " vs ", control_name)
  save_checked(plt, nrow(df), fn, width = 2.75, height = 2.9, layers = 1:2)
}
```

```{r fig.width=2.75, fig.height=2.9}
for (i in seq_len(nrow(protein_counts))) {
  plot_volcano(protein_counts$condition[i], protein_counts$control_condition[i])
}
```

## GSEA

Pre-ranked GSEA on GO biological process, one plot per insoluble/soluble fraction contrast.
Bar chart of NES coloured by -log10(FDR), following
`notebooks/01_whole_proteome/whole_proteome_visualization.Rmd` and
`notebooks/05_low_input_proteomics/low_input.Rmd`.

**Positive NES = enriched among proteins gaining intensity in the insoluble fraction.** The
top terms are shown per direction, so the insoluble-up side is always represented even when
the depleted side has the larger absolute scores.

```{r}
clean_term <- function(term) {
  str_split_fixed(term, "_", 2)[2] -> term
  gsub("\\b([a-z])", "\\U\\1", tolower(term), perl = TRUE) -> term
  str_replace_all(term, "_", " ") -> term
  str_wrap(term, 40) -> term
  return(term)
}

plot_gsea <- function(state, n_each = 10, fdr_cutoff = 0.01, dir = gsea_dir) {
  paste0(dir, state, "/gseapy.gene_set.prerank.report.csv") %>%
    read_csv(show_col_types = FALSE) %>%
    mutate(
      Term = lapply(Term, FUN = clean_term) %>% unlist(),
      `-log10(FDR)` = -log10(`FDR q-val`)
    ) %>%
    arrange(abs(NES)) -> df

  # Infinite -log10(FDR) is capped at the top of the fill scale, per repo convention.
  df$`-log10(FDR)`[is.infinite(df$`-log10(FDR)`)] <- 5.5

  sig <- df %>% filter(`FDR q-val` < fdr_cutoff)
  down_df <- sig %>% filter(NES < 0) %>% tail(n_each)
  up_df   <- sig %>% filter(NES > 0) %>% tail(n_each)
  plot_df <- rbind(down_df, up_df)

  plt <- plot_df %>%
    ggplot(aes(x = NES, y = reorder(Term, NES), fill = `-log10(FDR)`)) +
    geom_col(color = "black", linewidth = LINE_WIDTH, width = 0.5) +
    geom_vline(xintercept = 0, color = "black", linewidth = LINE_WIDTH) +
    my_theme() +
    labs(
      # bquote, not paste0: `"log"[2]` inside paste0 subscripts a character vector,
      # which yields NA rather than a subscript.
      x = bquote(.(state) ~ "log"[2] * "(Insoluble/Soluble) NES"),
      y = NULL,
      fill = expression("-log"[10] * "(FDR)")
    ) +
    theme(
      axis.text.y = element_text(size = 5, family = FONT_FAMILY),
      legend.text = element_text(margin = margin(l = 1)),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(4, "mm"),
      panel.grid.major = element_line(color = "grey90", linewidth = LINE_WIDTH)
    ) +
    scale_fill_continuous(low = "#B3CDE3", high = "#B10F7C", limits = c(1.5, 5.5))

  # One bar per selected term, so nothing is dropped by the fill scale limits.
  save_checked(plt, nrow(plot_df), paste0(dir, state, "/top_enriched_sets"),
               width = 3.5, height = 4)
  message(state, ": ", nrow(up_df), " up / ", nrow(down_df), " down of ",
          nrow(sig), " sets at FDR < ", fdr_cutoff)
  invisible(plot_df)
}
```

```{r fig.width=3.5, fig.height=4}
for (state in c("D2", "D8A", "D8C")) {
  plot_gsea(state)
}
```

## Verification

Every plotted point count must equal the tested count from the Python audit, and the totals
must reconcile against the 5330 input proteins.

```{r}
stopifnot(all(
  protein_counts$n_tested +
    protein_counts$n_dropped_group_too_small +
    protein_counts$n_dropped_other == protein_counts$n_input
))
stopifnot(nrow(volcano_data) == sum(protein_counts$n_tested))

# No point may fall outside its family's limits.
volcano_data %>%
  left_join(family_limits, by = "family") %>%
  summarise(
    outside_x = sum(log2_FC < x_lo | log2_FC > x_hi),
    outside_y = sum(neg_log10_pval < y_lo | neg_log10_pval > y_hi)
  ) -> outside
stopifnot(outside$outside_x == 0, outside$outside_y == 0)

protein_counts %>%
  select(condition, control_condition, n_input, n_tested,
         n_significant_up, n_significant_down, median_log2_FC)
```

# Revision run: HK3-129 + HK3-130

Everything below is the two-donor revision run, analysed and written by the revision section
of `solubility_analysis.ipynb`. The design differs from HK3-124 in three ways that matter to
the figures:

- **Two donors.** `d1` = HK3-129, `d2` = HK3-130, each contributing two TMT channel
  duplicates, so a state x fraction group has four channels. The replicate unit is the donor.
- **No normalization.** No soluble/insoluble protein concentrations were recorded for these
  runs, so there is no Final dilution factor to divide out. The fraction contrasts keep their
  large negative global offset; read them as a global shift, not as regulation.
- **Two-donor filter.** Only proteins quantified in both donors in all six groups are kept,
  4823 of 5569. All seven proteins of interest survive, as do all eight of the barplot panel.

```{r}
rev_results_dir <- paste0(data_dir, "revision/03_results/")
rev_volcano_dir <- paste0(rev_results_dir, "volcano_plots/")
rev_scatter_dir <- paste0(rev_results_dir, "scatterplots/")
rev_barplot_dir <- paste0(rev_results_dir, "barplots/")
rev_gsea_dir    <- paste0(rev_results_dir, "gsea/")

for (d in c(paste0(rev_volcano_dir, "figures"), paste0(rev_scatter_dir, "figures"),
            paste0(rev_barplot_dir, "figures"))) {
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
}

PROTEINS_OF_INTEREST <- c("MAP2K3", "MAP2K4", "LONP1", "HSPA9",
                          "DNAJA3", "NDUFA9", "VDAC1")

# The barplot panel, mirroring `BARPLOT_PROTEIN_SETS` in solubility.py: `PROTEINS_OF_INTEREST`
# plus PSMC5. GZMB is intentionally excluded from this panel (it's drawn separately below).
BARPLOT_PANEL <- c("MAP2K3", "MAP2K4", "LONP1", "HSPA9", "PSMC5",
                   "DNAJA3", "NDUFA9", "VDAC1")

# Positive controls: proteins with well-characterised solubility behaviour, drawn as their
# own panel rather than folded into the manuscript's eight.
POSITIVE_CONTROLS <- c("GZMB", "HSP90B1", "PRF1")

# Shared donor shape scale: circle = d1, square = d2, everywhere in this section.
donor_shapes <- c("d1" = 21, "d2" = 22)
```

## PCA

```{r fig.width=3.3, fig.height=2}
plot_revision_pca <- function(x = "PC1", y = "PC2") {
  pca_dir <- paste0(rev_results_dir, "pca_unnormalized/")
  percent <- read_csv(paste0(pca_dir, "percent_explained.csv"), show_col_types = FALSE)
  pcs <- split(percent$percent_explained, percent$principal_component)

  scores <- read_csv(paste0(pca_dir, "pca_results.csv"), show_col_types = FALSE) %>%
    mutate(
      condition = factor(condition, levels = group_levels),
      # `<state>-<fraction>_<donor>_<duplicate>`
      donor = str_split_fixed(channel_name, "_", 3)[, 2]
    )
  stopifnot(all(scores$donor %in% names(donor_shapes)))

  pad <- function(v) {
    r <- range(v)
    r + c(-1, 1) * diff(r) * 0.12
  }

  plt <- scores %>%
    ggplot(aes(x = .data[[x]], y = .data[[y]], fill = condition, shape = donor)) +
    geom_point(size = POINT_SIZE, alpha = 0.9, col = "black", stroke = POINT_STROKE) +
    scale_fill_manual(values = solubility_fills, drop = FALSE) +
    scale_shape_manual(values = donor_shapes, drop = FALSE) +
    coord_cartesian(xlim = pad(scores[[x]]), ylim = pad(scores[[y]])) +
    labs(x = paste0(x, " (", pcs[[x]], "%)"),
         y = paste0(y, " (", pcs[[y]], "%)")) +
    my_theme() +
    theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.position = "right",
      legend.key = element_blank(),
      legend.key.size = unit(0.25, "cm"),
      legend.margin = margin(l = -4),
      legend.text = element_text(size = FONT_SIZE, colour = "black",
                                 family = FONT_FAMILY, margin = margin(l = 1))
    ) +
    guides(fill = guide_legend(override.aes = list(size = 1.6, shape = 21)),
           shape = guide_legend(override.aes = list(size = 1.6)))

  save_checked(plt, nrow(scores), paste0(pca_dir, "pca_plot"), width = 3.3, height = 2)
}

plot_revision_pca()
```

## Volcano plots

Same nine comparisons as HK3-124, now on four channels per group (two donors x two TMT
duplicates), with the seven proteins of interest drawn on top and labelled.

```{r}
rev_volcano_data <- read_csv(paste0(rev_volcano_dir, "long_volcano_df.csv"),
                             show_col_types = FALSE)
rev_protein_counts <- read_csv(paste0(rev_volcano_dir, "volcano_protein_counts.csv"),
                               show_col_types = FALSE)

rev_family_limits <- rev_volcano_data %>%
  group_by(family) %>%
  summarise(
    x_lo = min(log2_FC), x_hi = max(log2_FC),
    y_lo = 0,            y_hi = max(neg_log10_pval),
    .groups = "drop"
  ) %>%
  mutate(
    x_pad = (x_hi - x_lo) * 0.05,
    y_pad = y_hi * 0.08,
    x_lo = x_lo - x_pad, x_hi = x_hi + x_pad,
    y_hi = y_hi + y_pad
  )

rev_family_limits
```

```{r fig.width=2.75, fig.height=2.9}
for (i in seq_len(nrow(rev_protein_counts))) {
  plot_volcano(
    rev_protein_counts$condition[i], rev_protein_counts$control_condition[i],
    data = rev_volcano_data, counts = rev_protein_counts,
    limits = rev_family_limits, out_dir = rev_volcano_dir,
    label_proteins = PROTEINS_OF_INTEREST
  )
}
```

## Scatterplots

log2FC against log2FC, styled after `plot_concordance` in
`notebooks/07_spin_reactivity/spin_reactivity_visualization.Rmd`: one symmetric range shared
by both axes so the identity line renders at 45 degrees, the grey identity line, and the
dashed diagonals one `BAND_CUTOFF` either side of it marking a **2-fold** difference between
the two contrasts. Ranges come from `coord_cartesian()`, never `scale_*_continuous(limits =)`.

`BAND_CUTOFF` is not `FC_CUTOFF`: the latter is the project's log2(1.5) significance cutoff,
which the volcanoes above use. The band is a stricter 2-fold, because `delta` is a difference of
two log2FCs and carries no p-value of its own - the band is the only thing separating a real
divergence from the scatter of two noisy measurements. It is read back from the Python layer,
which classified `concordance` with the same number.

The **state vs state** family - the fraction contrast log2(Insoluble/Soluble) of one state
against another - is what the percent-insoluble panels below are drawn from. A protein above
the band moves further into the insoluble fraction in the state on the y axis.

```{r}
scatter_summary <- read_csv(paste0(rev_scatter_dir, "scatter_summary.csv"),
                            show_col_types = FALSE)

# Read back from Python so the drawn band and its counts stay in step; not `FC_CUTOFF`.
stopifnot(length(unique(scatter_summary$fc_cutoff)) == 1)
BAND_CUTOFF <- scatter_summary$fc_cutoff[1]
BAND_FOLD <- 2 ^ BAND_CUTOFF
stopifnot(abs(BAND_FOLD - 2) < 1e-9)
scatter_summary %>%
  select(kind, x_label, y_label, n_joined, pearson_r, pearson_r_percent, median_delta,
         n_concordant, n_higher_on_x, n_higher_on_y)
```

`percent_insoluble/` holds the percent version of the state-vs-state family below.

```{r}
scatter_figure_path <- function(kind, fn_stem) {
  dir <- paste0(rev_scatter_dir, "figures/", kind, "/")
  dir.create(dir, recursive = TRUE, showWarnings = FALSE)
  paste0(dir, fn_stem)
}
```

### The same comparison in % insoluble

A state's soluble and insoluble shares sum to 100% by construction, so
log2(Insoluble/Soluble) is only a reparameterisation of "% insoluble" - the state-vs-state
family above and the barplot below already report one quantity on two scales. These panels
draw that family on the barplot's scale, so the two figures can be read against each other
directly.

Nothing about the classification changes. On percent axes the `|delta| > BAND_CUTOFF` band is
the odds curve

```
y / (100 - y) = 2 * x / (100 - x)
```

which the Python layer asserts the percent columns reproduce exactly, so the off-band protein
set and every count is the same as in the log2 panels - only the axes differ. The band bends,
so it is `geom_line` over a grid and not `geom_abline`; the identity line comes off the same
grid to keep the two in one place.

The axis is linear 0-100, the same scale the barplot uses. 77% of the 4823 proteins sit below
20% insoluble, so the cloud is dense in the bottom-left corner - that density is the finding,
not a plotting artefact, and reading it off the barplot's own axis is the point of the panel.

The percent is the insoluble share of the *channel means*, while a barplot bar is the mean of
the *per-channel shares*; `check_percent_agreement` reconciles the two (median 0.04 pp apart,
worst 6.8 pp).

```{r}
# The 2-fold band on percent axes, and the identity line, as curves. `band_grid` stops
# short of 0 and 100 because the odds are undefined at both ends.
FC_RATIO <- BAND_FOLD
band_grid <- tibble(x = seq(0.05, 99.95, length.out = 2000)) %>%
  mutate(
    odds = x / (100 - x),
    up   = 100 * (FC_RATIO * odds) / (1 + FC_RATIO * odds),
    down = 100 * (odds / FC_RATIO) / (1 + odds / FC_RATIO)
  )

plot_percent_insoluble_scatter <- function(x_label_in, y_label_in) {
  info <- scatter_summary %>% filter(x_label == x_label_in, y_label == y_label_in)
  stopifnot(nrow(info) == 1, info$kind == "state_vs_state")

  concordance_cols <- setNames(
    c("gray88", SIG_UP_COL, SIG_DOWN_COL),
    c("Concordant", paste("Higher in", info$y_short), paste("Higher in", info$x_short))
  )

  fn_stem <- paste0(gsub("/", "_over_", x_label_in), " vs ",
                    gsub("/", "_over_", y_label_in))
  df <- read_csv(paste0(rev_scatter_dir, fn_stem, ".csv"), show_col_types = FALSE) %>%
    mutate(concordance = factor(concordance, levels = names(concordance_cols)))
  stopifnot(nrow(df) == info$n_joined,
            !any(is.na(df$percent_insoluble_x)), !any(is.na(df$percent_insoluble_y)))

  # One span for both axes so identity renders at 45 degrees, padded 6%.
  span <- range(c(df$percent_insoluble_x, df$percent_insoluble_y))
  span <- span + c(-1, 1) * diff(span) * 0.06
  at <- function(f) span[1] + diff(span) * f

  counts <- tibble(
    x = c(at(0.04), at(0.04), at(0.96)),
    y = c(at(0.96), at(0.88), at(0.04)),
    label = c(paste0("R = ", sprintf("%.2f", info$pearson_r_percent)),
              paste0(info$n_higher_on_y, " higher in ", info$y_short),
              paste0(info$n_higher_on_x, " higher in ", info$x_short)),
    hjust = c(0, 0, 1)
  )

  poi <- df %>% filter(protein_of_interest)
  stopifnot(nrow(poi) == info$n_proteins_of_interest)

  set.seed(12345)
  plt <- ggplot(df, aes(x = percent_insoluble_x, y = percent_insoluble_y,
                        fill = concordance)) +
    geom_line(data = band_grid, aes(x = x, y = x), inherit.aes = FALSE,
              linewidth = LINE_WIDTH, colour = UNCHANGED_COL) +
    geom_line(data = band_grid, aes(x = x, y = up), inherit.aes = FALSE,
              linetype = "dashed", linewidth = LINE_WIDTH, colour = "black") +
    geom_line(data = band_grid, aes(x = x, y = down), inherit.aes = FALSE,
              linetype = "dashed", linewidth = LINE_WIDTH, colour = "black") +
    geom_point(data = df %>% filter(concordance == "Concordant"),
               shape = 21, size = POINT_SIZE - 1.5, stroke = POINT_STROKE / 2,
               colour = "gray88", alpha = 0.4) +
    geom_point(data = df %>% filter(concordance != "Concordant"),
               shape = 21, size = POINT_SIZE - 1.5, stroke = POINT_STROKE / 2,
               colour = "gray55", alpha = 0.8) +
    geom_text(data = counts, aes(x = x, y = y, label = label, hjust = hjust),
              inherit.aes = FALSE, size = FONT_SIZE_MM, colour = "black") +
    geom_point(data = poi, shape = 21, size = POINT_SIZE - 1.5,
               stroke = POINT_STROKE / 2, colour = "black", alpha = 1) +
    geom_text_repel(
      data = poi, aes(x = percent_insoluble_x, y = percent_insoluble_y, label = protein),
      inherit.aes = FALSE, size = FONT_SIZE_MM, colour = "black",
      box.padding = 0.4, min.segment.length = 0, segment.size = LINE_WIDTH) +
    scale_fill_manual(values = concordance_cols, drop = FALSE, name = NULL) +
    coord_cartesian(xlim = span, ylim = span) +
    labs(
      x = paste0("% insoluble, ", info$x_short),
      y = paste0("% insoluble, ", info$y_short)
    ) +
    my_theme() +
    theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.position = "top",
      legend.key = element_blank(),
      legend.margin = margin(l = -5, b = -5.5),
      legend.box.margin = margin(l = -5, b = -0.8),
      legend.key.spacing.x = unit(-0.05, "cm")
    ) +
    scale_x_continuous(breaks = seq(0, 100, 25)) +
    scale_y_continuous(breaks = seq(0, 100, 25)) +
    guides(fill = guide_legend(nrow = 1, override.aes = list(size = 1.6, alpha = 1)))

  # Layers 1-3 are the identity line and the band, 4-5 the cloud, 6 the annotations,
  # 7-8 the highlight. The folder carries the "percent insoluble" half of the name.
  save_checked(plt, nrow(df), scatter_figure_path("percent_insoluble", fn_stem),
               width = 3, height = 3.4, layers = 4:5)
}
```

```{r fig.width=3, fig.height=3.4}
state_pairs <- scatter_summary %>% filter(kind == "state_vs_state")
for (i in seq_len(nrow(state_pairs))) {
  plot_percent_insoluble_scatter(state_pairs$x_label[i], state_pairs$y_label[i])
}
```

### Cysteine reactivity changes

The same percent-insoluble family once more, this time coloured by whether a protein carries a
cysteine reactivity change in `notebooks/03_reactivity`. The question is whether a protein whose
cysteines changed reactivity also moved between the detergent fractions - the only panel in this
document that reads another assay.

Two categories, assigned in `solubility.py` (`load_reactivity_changes`) and carried into the
scatter CSVs as `reactivity_change` / `reactivity_label`, so the R layer never restates the
mapping:

- **Reactivity change** - the union across D4A/D4C/D8A/D8C, 180 proteins, 136 detected here.
  Unioned rather than state-matched: this run has no D4 states to match against, and a per-panel
  highlight set would change the coloured points from panel to panel.
- **Other** - the grey cloud.

Labels come from `reactivity_named`: `REACTIVITY_NAMED_PROTEINS` in `solubility.py` - MAP2K3,
MAP2K4, LONP1, HSPA9, PSMC5, DNAJA3, NDUFA9, VDAC1, eight in all. The named set is **not** a
subset of the coloured category - some of the eight carry a reactivity change and some do not -
so these labels are drawn in black rather than in the category colour, which would leave the
rest grey on grey.

```{r}
# One viridis sample from the repo's categorical palette.
reactivity_cols <- c(
  "Reactivity change" = "#440154",
  "Other"             = "gray88"
)

# One alpha for every point in this panel, highlighted or not. A highlighted point is a
# member of the same cloud, so it is drawn at the cloud's opacity and picked out by its fill
# and outline rather than by being the only solid mark on the panel.
CLOUD_ALPHA <- 0.4

plot_percent_insoluble_reactivity <- function(x_label_in, y_label_in) {
  info <- scatter_summary %>% filter(x_label == x_label_in, y_label == y_label_in)
  stopifnot(nrow(info) == 1, info$kind == "state_vs_state")

  fn_stem <- paste0(gsub("/", "_over_", x_label_in), " vs ",
                    gsub("/", "_over_", y_label_in))
  df <- read_csv(paste0(rev_scatter_dir, fn_stem, ".csv"), show_col_types = FALSE) %>%
    mutate(reactivity_label = factor(reactivity_label, levels = names(reactivity_cols)))
  stopifnot(
    nrow(df) == info$n_joined,
    !any(is.na(df$reactivity_label)),
    sum(df$reactivity_change) == info$n_reactivity_change,
    sum(df$reactivity_named) == info$n_reactivity_named,
    # The two categories partition the table; if they ever stopped doing so the layer split
    # below would quietly draw a protein twice or not at all.
    sum(df$reactivity_change) + sum(df$reactivity_label == "Other") == nrow(df)
  )

  span <- range(c(df$percent_insoluble_x, df$percent_insoluble_y))
  span <- span + c(-1, 1) * diff(span) * 0.06
  at <- function(f) span[1] + diff(span) * f

  changed <- df %>% filter(reactivity_change)
  named <- df %>% filter(reactivity_named)

  # Both annotations stack in the top-left corner. The highlighted points here sit low on
  # both axes, so a bottom-right count is repelled into by their labels.
  annot <- tibble(
    x = at(0.04),
    y = c(at(0.96), at(0.89)),
    hjust = 0,
    label = c(paste0("R = ", sprintf("%.2f", info$pearson_r_percent)),
              paste0(info$n_reactivity_change, " of ", info$n_reactivity_source,
                     " reactivity-change proteins detected"))
  )

  set.seed(12345)
  plt <- ggplot(df, aes(x = percent_insoluble_x, y = percent_insoluble_y,
                        fill = reactivity_label)) +
    geom_abline(slope = 1, intercept = 0, linewidth = LINE_WIDTH, colour = UNCHANGED_COL) +
    # Reactivity-change points are a member of this cloud, not a separate population - same
    # size/shape/alpha as the rest, just a black outline instead of gray55.
    geom_point(data = df %>% filter(reactivity_label == "Other"),
               shape = 21, size = POINT_SIZE - 1.5, stroke = POINT_STROKE / 2,
               colour = "gray55", alpha = CLOUD_ALPHA) +
    geom_point(data = changed, shape = 21, size = POINT_SIZE - 1.5,
               stroke = POINT_STROKE / 2, colour = "black", alpha = CLOUD_ALPHA) +
    geom_text(data = annot, aes(x = x, y = y, label = label, hjust = hjust),
              inherit.aes = FALSE, size = FONT_SIZE_MM, colour = "black") +
    # Labelled points redrawn opaque for a visible anchor - fill unchanged, adds nothing new
    # to the drawn-point audit.
    geom_point(data = named, shape = 21, size = POINT_SIZE - 1.5,
               stroke = POINT_STROKE / 2, colour = "black", alpha = 1) +
    geom_text_repel(
      data = named, aes(x = percent_insoluble_x, y = percent_insoluble_y, label = protein),
      inherit.aes = FALSE, size = FONT_SIZE_MM, colour = "black",
      box.padding = 0.4, min.segment.length = 0, segment.size = LINE_WIDTH,
      max.overlaps = Inf, seed = 12345) +
    scale_fill_manual(values = reactivity_cols, drop = FALSE, name = NULL) +
    coord_cartesian(xlim = span, ylim = span) +
    labs(
      x = paste0("% insoluble, ", info$x_short),
      y = paste0("% insoluble, ", info$y_short)
    ) +
    my_theme() +
    theme(
      aspect.ratio = 1,
      legend.title = element_blank(),
      legend.position = "top",
      legend.key = element_blank(),
      legend.margin = margin(l = -5, b = -5.5),
      legend.box.margin = margin(l = -5, b = -0.8),
      legend.key.spacing.x = unit(-0.05, "cm")
    ) +
    scale_x_continuous(breaks = seq(0, 100, 25)) +
    scale_y_continuous(breaks = seq(0, 100, 25)) +
    guides(fill = guide_legend(nrow = 1, override.aes = list(size = 1.6, alpha = 1)))

  # Cloud split across layers 2-3; audit sums both since together they're every protein.
  save_checked(plt, nrow(df), scatter_figure_path("percent_insoluble_reactivity", fn_stem),
               width = 2, height = 2.2, layers = 2:3)
}
```

```{r fig.width=2, fig.height=2.2}
for (i in seq_len(nrow(state_pairs))) {
  plot_percent_insoluble_reactivity(state_pairs$x_label[i], state_pairs$y_label[i])
}
```

A `notes.txt` beside the three panels, on the same footing as the complex notes above: written
from here so it regenerates with the figures, and duplicated into `README.md` because `data/` is
gitignored.

```{r}
reactivity_note <- r"(Cysteine reactivity changes
=================================================

WHAT IS COLOURED
136 of the 180 proteins carrying a cysteine reactivity change in notebooks/03_reactivity,
unioned across D4A/D4C/D8A/D8C, joined on gene symbol.
Labelled: MAP2K3, MAP2K4, LONP1, HSPA9, PSMC5, DNAJA3, NDUFA9, VDAC1 (REACTIVITY_NAMED_PROTEINS
in solubility.py), eight in all.

WHAT THE PANELS SHOW    (median percentage-point shift; background in brackets)
Band = a 2-fold difference between the two contrasts.
  D2  -> D8A   +2.21 pp   (+1.84)    14.0% off-band vs 11.8%   OR 1.22, p = 0.42
  D2  -> D8C   -0.25 pp   (-0.40)     7.4% off-band vs  6.5%   OR 1.14, p = 0.72
  D8A -> D8C   -1.99 pp   (-2.07)     5.9% off-band vs  7.9%   OR 0.73, p = 0.52

INTERPRETATION
Read this as a negative result. The reactivity-change group does not shift in a direction
of its own in any panel - its median percentage-point shift is within a quarter of a point
of the background in all three, and a Mann-Whitney on the shifts is far from significant
everywhere (p = 0.40, 0.69, 0.84). Nor is it enriched for proteins that leave the band at
all, in either direction: no panel comes near significance by Fisher, and the D8A -> D8C
panel points the other way.

A note on the band width, because it changed the conclusion. At the log2(1.5) band this
family used previously, D2 -> D8A did show a nominal enrichment for off-band proteins
(46.3% vs 36.2%, OR 1.52, p = 0.019). It does not survive the move to 2-fold (14.0% vs
11.8%, OR 1.22, p = 0.42), which is what one would expect of an effect carried by the
proteins sitting nearest the band edge. Do not quote the 1.5-fold number.

CAVEATS
The two assays measure different things on different peptides: a reactivity change is a
cysteine-level call from TMT-ABPP, solubility is a protein-level fraction ratio. A protein
can change reactivity at one cysteine with no change in where the bulk of it partitions,
and the panel would show exactly what it shows.
The union across conditions mixes D4 and D8 timepoints into a highlight set drawn on D2/D8
axes; a state-matched set was considered and rejected because it changes the coloured
points from panel to panel, but it does mean the highlight is not timepoint-specific.
Equal protein mass was labelled from each fraction and this run has no Final dilution
factor, so a shift in insoluble share is a position relative to the rest of the pellet,
not a literal claim about how much protein is insoluble.
)"

writeLines(reactivity_note,
           paste0(rev_scatter_dir, "figures/percent_insoluble_reactivity/notes.txt"))
cat("wrote percent_insoluble_reactivity/notes.txt\n")
```

## Proteins of interest

The manuscript's eight proteins: `PROTEINS_OF_INTEREST` plus PSMC5, one of the 19S regulatory
ATPase subunits.

Each state's soluble and insoluble share, which sum to 100% by construction. Every replicate
channel is plotted: four points per bar, drawn as plain circles - two donors x two TMT
duplicates. Soluble duplicate `i` is paired with insoluble duplicate `i` of the same donor,
the only pairing available across fractions.

The error bar is the SE of those four points. Two of every four are technical duplicates, so
it is an SE over channels, not over donors, and understates donor-to-donor variability; the
points are not shape-coded by donor, so that spread is not resolved visually here.

**This is a relative display.** Equal protein *mass* was labelled from each fraction and no
Final dilution factor exists for these runs, so the channel ratio reports a protein's share
within its own fraction. A 90% soluble bar does not mean 90% of the protein is soluble; it
means the protein sits that much higher in the soluble fraction's distribution than in the
insoluble fraction's. Compare bars across states, not against 50.

As with the scatterplots, the panels get their own folder: `both_fractions/` for the pair
of bars below.

**Positive controls.** `POSITIVE_CONTROLS` (GZMB, HSP90B1, PRF1) get their own panel rather
than joining the manuscript's eight - GZMB was on the label set before the revision, and
HSP90B1/PRF1 are ER- and secretory-granule-anchored proteins with well-characterised
solubility behaviour, so the panel is a sanity check rather than a manuscript figure. Same
construction as above, one row of three facets.

```{r}
barplot_figure_path <- function(kind, fn_stem) {
  dir <- paste0(rev_barplot_dir, "figures/", kind, "/")
  dir.create(dir, recursive = TRUE, showWarnings = FALSE)
  paste0(dir, fn_stem)
}
```

```{r}
# `name` is the key `fraction_shares` wrote the CSV under; `proteins` fixes the facet order.
# A panel that lost a subunit or a channel fails here rather than drawing short.
read_panel_shares <- function(name, proteins) {
  shares <- read_csv(paste0(rev_barplot_dir, "fraction_shares_", name, ".csv"),
                     show_col_types = FALSE) %>%
    mutate(
      group = factor(group, levels = group_levels),
      state = factor(state, levels = c("D2", "D8A", "D8C")),
      fraction = factor(fraction, levels = c("Soluble", "Insoluble")),
      protein = factor(protein, levels = proteins),
      donor = factor(donor, levels = names(donor_shapes))
    )

  stopifnot(!any(is.na(shares$group)), !any(is.na(shares$protein)),
            !any(is.na(shares$donor)))
  stopifnot(setequal(levels(droplevels(shares$protein)), proteins),
            nrow(shares) <= length(proteins) * 3 * 2 * 4)

  # TMT non-detections: a share is NaN in both fractions or neither (shared denominator), so
  # dropping them leaves whole pairs and `save_checked` gets an exact point count.
  missing <- shares %>% filter(is.na(percent))
  if (nrow(missing) > 0) {
    message(name, ": dropping ", nrow(missing), " non-detected channel(s): ",
            paste(missing$protein, missing$group, missing$replicate, collapse = "; "))
    shares <- shares %>% filter(!is.na(percent))
  }
  # Whatever was dropped, both donors must still stand behind every protein x group.
  stopifnot(nrow(count(shares, protein, group, donor)) == length(proteins) * 6 * 2)

  # Every protein x state x channel must total 100 across the two fractions.
  shares %>%
    group_by(protein, state, replicate) %>%
    summarise(total = sum(percent), .groups = "drop") %>%
    pull(total) %>%
    {stopifnot(all(abs(. - 100) < 1e-6))}

  shares
}

panel_stats <- function(shares) {
  shares %>%
    group_by(protein, state, fraction, group) %>%
    summarise(n = n(), mean = mean(percent), sd = sd(percent), .groups = "drop") %>%
    mutate(se = sd / sqrt(n))
}

insoluble_table <- function(stats) {
  stats %>% filter(fraction == "Insoluble") %>%
    select(protein, state, mean, se) %>%
    pivot_wider(names_from = state, values_from = c(mean, se)) %>%
    mutate(across(where(is.numeric), ~ round(.x, 1)))
}

poi_shares <- read_panel_shares("proteins_of_interest", BARPLOT_PANEL)
poi_stats <- panel_stats(poi_shares)

insoluble_table(poi_stats)

pc_shares <- read_panel_shares("positive_controls", POSITIVE_CONTROLS)
pc_stats <- panel_stats(pc_shares)

insoluble_table(pc_stats)
```

```{r fig.width=4, fig.height=2}
# ~0.8in per facet column plus ~0.7in for the right-hand legend, at `legend_theme()`'s key size.
plot_poi_barplot <- function(shares, stats, out_name = "proteins_of_interest",
                             width = 4, height = 2, nrow = 2) {
  dodge <- position_dodge(width = 0.8)
  set.seed(20260826)  # reproducible point jitter

  plt <- ggplot(shares, aes(x = state, y = percent, fill = group)) +
    # 0.6 against a 0.8 dodge, as in the dodged faceted bargraph of
    # notebooks/04_metabolomics - not the shared BARPLOT_WIDTH, which is 0.75 and leaves the
    # soluble and insoluble bars of a state touching.
    geom_bar(stat = "summary", fun = "mean", width = 0.6, alpha = 0.8,
             position = dodge, colour = "black", linewidth = LINE_WIDTH,
             show.legend = TRUE) +
    geom_errorbar(
      data = stats,
      mapping = aes(x = state, ymin = mean - se, ymax = mean + se, group = group),
      inherit.aes = FALSE, position = dodge, width = 0.3, linewidth = LINE_WIDTH) +
    # `group = group` pins points to their own bar - without it position_dodge splits each
    # bar's points into offset columns.
    geom_point(aes(group = group), shape = 21,
               size = POINT_SIZE - 1.6, stroke = POINT_STROKE / 2, colour = "black",
               position = position_jitterdodge(jitter.width = 0.35, jitter.height = 0,
                                               dodge.width = 0.8)) +
    scale_fill_manual(values = solubility_fills, drop = FALSE) +
    # Shares run 0-100 by construction, so a fixed range is safe; expand only upward for
    # the error bars.
    coord_cartesian(ylim = c(0, 105)) +
    scale_y_continuous(breaks = seq(0, 100, 25), expand = c(0, 0)) +
    facet_wrap(~protein, nrow = nrow) +
    labs(x = NULL, y = "% of state total") +
    my_theme() +
    theme(
      strip.background = element_blank(),
      strip.text = element_text(size = FONT_SIZE, family = FONT_FAMILY, colour = "black"),
      panel.spacing = unit(1, "mm"),
      panel.border = element_blank(),
      axis.line.x = AXIS_LINE,
      axis.line.y = AXIS_LINE,
      legend.title = element_blank(),
      legend.position = "right",
      legend.key = element_blank(),
      legend.key.size = unit(0.2, "cm"),
      legend.margin = margin(l = -5, b = -4.5),
      legend.box.margin = margin(l = -5, b = -5.7),
      # NOT `legend_theme()`'s `margin(l = -5)`: at this key size the swatch is only 5.3pt
      # wide and `l = -5` lands the text almost inside it.
      legend.text = element_text(size = FONT_SIZE, colour = "black",
                                 family = FONT_FAMILY, margin = margin(l = 1)),
      plot.margin = margin(t = 1, r = 1, b = 1, l = 1)
    ) +
    guides(fill = guide_legend(override.aes = list(shape = NA)))

  # Layer 3 is the channel points: one per row of `shares`, i.e. proteins x 3 states x 2
  # fractions x 2 donors x 2 TMT duplicates.
  save_checked(plt, nrow(shares), barplot_figure_path("both_fractions", out_name),
               width = width, height = height, layers = 3)
}

plot_poi_barplot(poi_shares, poi_stats, width = 4, height = 2, nrow = 2)
```

```{r fig.width=4, fig.height=1.07}
# height chosen so the single facet row here matches the main panel's per-row height exactly
# (measured from each plot's gtable: main panel's panel rows are `null`-unit and split 2 - 0.61
# fixed in = 0.69 in/row; this panel has only one strip row so its fixed component is smaller,
# 0.37 in - solving `0.69*1 + 0.37` gives 1.07, not half of the main panel's total height).
plot_poi_barplot(pc_shares, pc_stats, out_name = "positive_controls",
                 width = 4, height = 1.07, nrow = 1)
```

## GSEA

Same pre-ranked GO:BP analysis as HK3-124, on the three fraction contrasts of the revision
run. Positive NES = enriched among proteins gaining intensity in the insoluble fraction.

```{r fig.width=3.5, fig.height=4}
for (state in c("D2", "D8A", "D8C")) {
  plot_gsea(state, dir = rev_gsea_dir)
}
```

## Verification

```{r}
stopifnot(all(
  rev_protein_counts$n_tested +
    rev_protein_counts$n_dropped_group_too_small +
    rev_protein_counts$n_dropped_other == rev_protein_counts$n_input
))
stopifnot(nrow(rev_volcano_data) == sum(rev_protein_counts$n_tested))

rev_volcano_data %>%
  left_join(rev_family_limits, by = "family") %>%
  summarise(
    outside_x = sum(log2_FC < x_lo | log2_FC > x_hi),
    outside_y = sum(neg_log10_pval < y_lo | neg_log10_pval > y_hi)
  ) -> rev_outside
stopifnot(rev_outside$outside_x == 0, rev_outside$outside_y == 0)

# All seven proteins of interest must survive the two-donor filter and appear in every
# comparison.
poi_present <- rev_volcano_data %>%
  filter(protein %in% PROTEINS_OF_INTEREST) %>%
  count(condition, control_condition)
stopifnot(all(poi_present$n == length(PROTEINS_OF_INTEREST)))

rev_protein_counts %>%
  select(condition, control_condition, n_input, n_tested,
         n_significant_up, n_significant_down, median_log2_FC)
```
