---
title: "exhaustion_reactivity"
author: "Henry Sanford"
date: "2024-07-09"
output: html_document
---

## Setup

```{r, results}
# set working directory to script location

# load figure settings
source("../../bin/figure_utils.R")

results_dir <- "../../data/reactivity/results"

protein_list_dir <- "/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Exhaustion manuscript/02_Mass-spectrometry data/Reference lists/"

add_back_dir <- "../data/reactivity/atp_add_back"

```

## Waterfall plots

```{r, fig.height=3, fig.width=2}
reactivity_rank_dir <- paste0(results_dir,"02_cysteine_reactivity_rank/")
for (condition in conditions){
  paste0(reactivity_rank_dir,condition,"_cysteine_reactivity_rank.csv") -> fn
  read_csv(fn, show_col_types = FALSE) -> df
  df %>% mutate(residue_rank = rank(-LFC_tmt_abpp)) %>% 
    mutate(`Cysteine reactivity in D8C` = case_when(LFC_tmt_abpp > 1 ~ "Higher",
                                LFC_tmt_abpp < -1 ~ "Lower",
                                TRUE ~ "Unchanged")) %>%  
    mutate(name = paste0(protein, "_", residue))-> df
  df %>% write_csv(fn)
  
  df %>% filter(protein %in% 
                  c("PIP5K1B", 
                    "PIP5K1A", 
                    "PIP5K1C",
                    "PIP4K2A",
                    "PIP4K2B",
                    "PIP4K2C",
                    "PIK3CB",
                    "PIP4P1",
                    "PLCB1",
                    "INPP5E",
                    "PTEN"
                    )) -> label_data
  
  df %>% ggplot(aes(x=residue_rank, y = LFC_tmt_abpp, color = `Cysteine reactivity in D8C`)) + 
    geom_point(aes(fill= `Cysteine reactivity in D8C`), 
               size=POINT_SIZE - 1, 
               shape=21, 
               color = "transparent",
                stroke=POINT_STROKE / 4,
               show.legend = FALSE) + 
    geom_hline(yintercept = c(1,-1), linetype="dashed", size=LINE_WIDTH, color = "black") + 
    labs(
      x = "residue rank",
      y = substitute(paste('cysteine reactivity, log'[2],
                           "(",n, "/D2)"), list(n = condition))
    ) +
    geom_text_repel(data = label_data, 
                     size=FONT_SIZE_MM,
                     aes(label=name), 
                    color = "black",
                     min.segment.length = unit(0, 'lines'), 
                    segment.size = LINE_WIDTH,
                     box.padding = 1,
                     show.legend = FALSE,
                     family="Arial") + 
    geom_point(data=label_data, stroke=POINT_STROKE, shape=21, color="black", size=POINT_SIZE - 1,
               show.legend = FALSE) +
    scale_fill_manual(values = regulation_colors, name = paste0("Cysteine reactivity in ", condition)) + 
    ylim(-4.5,6) +
    my_theme() + 
    theme(legend.position = "top",
      legend.box.background = element_blank(),
      aspect.ratio=1,
      legend.key = element_blank(),
      legend.margin = margin(l = -5, b = -5.5),
      legend.box.margin = margin(l = -5, b = -5.7),
      legend.key.spacing.y = unit(-1, "cm"),
      legend.key.spacing.x = unit(-0.05, "cm"))-> plot
  print(plot)
  save_plot(paste0(reactivity_rank_dir, condition, "_cysteine_reactivity_rank"), width = 4, height = 2)
}
```


## WP RC one peptide scatterplot

```{r, fig.width=2, fig.height=2}
plot_lim = 3.8 + 0.07
nudge_val = 2
reactivity_vs_wp_dir <- paste0(results_dir, "03_reactivity_vs_wp")
line_df <- data.frame(x1 = c(-plot_lim, -plot_lim + 1), 
                      x2 = c(plot_lim - 1, plot_lim), 
                      y1 = c(-plot_lim + 1, -plot_lim), 
                      y2 = c(plot_lim, plot_lim - 1))

one_peptide_scatterplot <- function(fil_condition, highlights){
  
  paste0(reactivity_vs_wp_dir, "/",fil_condition, "_D2_reactivity_wp_comparison_data.csv") %>%
  read_csv(show_col_types = FALSE) -> full_df
  
  full_df %>% mutate(
    cysteine_reactivity_regulation = case_when(
      ((tmt_abpp / whole_proteome) < 0.5) ~ "Lower",
      ((tmt_abpp / whole_proteome) > 2) ~ "Higher",
      TRUE ~ "Unchanged"
    )
  ) %>% 
    mutate(LFC_wp = log2(whole_proteome / 100)) %>% 
    filter(is.finite(LFC_wp)) %>%
    filter(num_quantified_peptides == 1) %>%
    filter(condition == fil_condition)  -> full_df
  
  full_df %>% 
    mutate(name = paste0(protein, "_", residue)) -> df
  
  full_df %>% write_csv(
    paste0(reactivity_vs_wp_dir, "/",fil_condition, "_D2_reactivity_wp_comparison_data_1_peptide.csv")
  )
  
  df %>% 
    filter(protein %in% highlights) -> label_data
  
  df %>% 
    ggplot(aes(x=LFC_tmt_abpp, 
               y=LFC_wp, 
               fill=cysteine_reactivity_regulation)) + 
    geom_point(
               color="black" ,
               size = 1,
               stroke=POINT_STROKE / 2,
               shape=21, 
               alpha=0.6,
               show.legend = FALSE) + 
    geom_point(data=label_data, 
               color="black" ,
               size = 1,
               stroke=POINT_STROKE / 2,
               shape=21, 
               show.legend = FALSE) + 
    scale_fill_manual(values = regulation_colors) +
    scale_x_continuous(breaks = c(-4,-3,-2,-1,0,1,2,3,4),
                       limits = c(-plot_lim, plot_lim),
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(-4,-3,-2,-1,0,1,2,3,4),
                       limits = c(-plot_lim, plot_lim),
                       expand = c(0,0)) +
   geom_text_repel(data = label_data %>% filter(cysteine_reactivity_regulation == "Lower"), 
                     aes(label=name, color = cysteine_reactivity_regulation), 
                     size=FONT_SIZE_MM,
                    segment.size = LINE_WIDTH,
                     min.segment.length = unit(0, 'lines'), 
                     nudge_y=nudge_val, 
                   ylim = c(0, NA),
                     show.legend = FALSE) + 
    geom_text_repel(data = label_data %>% filter(cysteine_reactivity_regulation == "Higher"), 
                     aes(label=name, color = cysteine_reactivity_regulation), 
                     size=FONT_SIZE_MM,
                    segment.size = LINE_WIDTH,
                     min.segment.length = unit(0, 'lines'), 
                     nudge_y=-nudge_val, 
                   ylim = c(NA, 0),
                     show.legend = FALSE) + 
    scale_color_manual(values = regulation_colors) +
    labs(
      x = substitute(paste('cysteine reactivity, log'[2],
                           "(",n, "/D2)"), list(n = fil_condition)),
      y = substitute(paste('protein expression, log'[2],
                           "(",n, "/D2)"), list(n = fil_condition)),
    ) +
    geom_segment(data = line_df,
      aes(x = x1, y = y1, 
          xend = x2, yend = y2),
      inherit.aes = FALSE,
      linetype = "dashed",
      size = LINE_WIDTH,
      color = "black") +
    my_theme() +
    theme(aspect.ratio=1, 
          legend.key.spacing.y = unit(-2,"mm"),
          legend.key = element_blank(),)-> plot
    print(plot)
    save_plot(
      paste0(results_dir, 
             "03_reactivity_vs_wp/", 
             fil_condition,"_D2_scatter"), 
      width=2, height=2)
}


one_peptide_scatterplot(
  "D8C",
  c("HTRA2","RAB39B","PYCARD","TFRC","SDCBP","NDUFB7","TIMM10")
)

one_peptide_scatterplot(
  "D8A",
  c("HTRA2","RAB39B","TFRC")
)

for(condition in c("D4A", "D4C")){
 one_peptide_scatterplot(
  condition,
  c()) 
}
```

## Reactivity change dotplots

```{r, fig.height = 4, fig.width=2}
lbl_nudge <- 0.4

fil_condition <- "D8C"
rc_dotplot <- function(targets,
                       title,
                       ratio_aspect = 1,
                       ht = 4,
                       wd = 2,
                       y_max = 3,
                       y_min = -3) {
  plt_df <- paste0(reactivity_vs_wp_dir, "/rc_df_long_format.csv") %>%
    read.csv() %>%
    filter(condition == fil_condition) %>%
    filter(protein %in% targets) %>%
    mutate(
      `Cysteine reactivity in D8C` = case_when(
        reactivity_change != "True" ~ "Unchanged",
        LFC_tmt_abpp > 0 ~ "Higher",
        LFC_tmt_abpp < 0 ~ "Lower",
        TRUE ~ "Unchanged"
      ),
      protein = factor(protein, levels = targets)
    )
  
  lbl_data <- plt_df %>%
    filter(reactivity_change == "True")
  
  wp_data <- plt_df %>% 
    select(uniprot, protein, condition, LFC_WP) %>%
    mutate(`Cysteine reactivity in D8C` = "Protein expression") %>% 
    distinct()

  plt_df %>%
    arrange(by = "protein") %>%
    ggplot(aes(x = protein, y = LFC_tmt_abpp, fill = `Cysteine reactivity in D8C`)) +
    geom_hline(yintercept = c(1,-1),
               linetype = "dashed",
               size = LINE_WIDTH) +
    # cysteine data
    geom_point(
      size = POINT_SIZE - 0.75,
      shape = 21,
      color = "black",
      alpha = 0.9,
      stroke = POINT_STROKE
    ) +
    # residue labels
    geom_text(
      data = lbl_data,
      mapping = aes(
        y = LFC_tmt_abpp + case_when(
          LFC_tmt_abpp > 0 ~ lbl_nudge,
          LFC_tmt_abpp < 0 ~ -lbl_nudge,
          TRUE ~ lbl_nudge
        ),
        color = `Cysteine reactivity in D8C`,
        label = residue
      ),
      size = FONT_SIZE,
      size.unit = "pt",
      show.legend = FALSE
    ) +
    # whole proteome data (triangle)
    geom_point(
      data = wp_data,
      aes(x = protein, y = LFC_WP),
      shape = 24,
      color = "black",
      size = POINT_SIZE - 0.75,
      stroke = POINT_STROKE,
      alpha = 0.9
    ) +
    # aesthetics
    my_theme() +
    scale_fill_manual(values = regulation_colors,
                      breaks = names(regulation_colors)) +
    scale_color_manual(values = regulation_colors,
                       breaks = names(regulation_colors)) +
    scale_y_continuous(breaks = round(y_min):round(y_max),
                       limits = c(y_min, y_max)) +
    guides(fill = guide_legend(byrow = TRUE,
                               nrow = TRUE)) +
    theme(
      axis.text.x = element_text(
        color = "black",
        angle = 45,
        size = FONT_SIZE,
        family = "Arial",
        vjust = 1,
        hjust = 1
      ),
      legend.key.spacing.y = unit(-2.5, "mm"),
      legend.margin = margin(b = -15),
      legend.position = "top",
      #aspect.ratio = ratio_aspect,
      legend.title = element_text(margin = margin(b = -2)),
      legend.key = element_blank(),
      legend.title.position = "top",
      legend.key.spacing = unit(0, "mm"),
      legend.spacing.y = unit(-0, "mm"),
      legend.spacing = unit(0, "mm")
    ) +
    labs(y = substitute(
      paste('cysteine reactivity, log'[2],
            "(", n, "/D2)"),
      list(n = fil_condition)
    ),
    x = "") +
    guides(fill = guide_legend(ncol = 2)) -> plot
  print(plot)
  save_plot(
    paste0(results_dir, "06_dotplots/", title),
    height = ht,
    width = wd
  )
}
```




### Multiple pathways

```{r, fig.height=2.63, fig.width=4.6}
rc_dotplot(
  c(
    "CDC123",
    "TUFM", 
    "TLN1", 
    "LIMA1", 
    "PLS3",
    "AGPS", 
    "ACSL4",
    "SACS", 
    "BAG5",
    "DNAJC11", 
    "LONP1", 
    "HSPA9", 
    "HSPH1",
    "GFER",
    "NDUFV1",
    "MAP2K3",
    "MAP2K4"
    ),
  "S3D D8C vs D2",
  ratio_aspect = 0.5,
  wd = 4.6,
  ht = 2.63,
  y_min = -3,
  y_max = 3.5
)
```


```{r, fig.width = 2.88, fig.height = 2.6861}
facetted_multi_condition_dotplot <-
  function(targets,
           display_conditions,
           title,
           additional_cysteines_to_label = c(),
           fig_height,
           fig_width,
           y_min = -4.5,
           y_max = 2) {
    paste0(reactivity_vs_wp_dir, "/rc_df_long_format.csv") %>% 
      read_csv(show_col_types = FALSE) %>%
      filter(protein %in% targets,
             condition %in% display_conditions) %>%
      mutate(`PPAT reactivity in D8C` = case_when(
        (reactivity_change == TRUE) & (LFC_tmt_abpp > 0) ~ "Higher",
        (reactivity_change == TRUE) & (LFC_tmt_abpp < 0) ~ "Lower",
        TRUE ~ "Unchanged"
      ),
      condition = factor(condition, levels = names(cols)),
      protein = factor(protein, levels = targets)) -> plt_df
    
    plt_df %>%
      filter(reactivity_change == TRUE |
               residue %in% additional_cysteines_to_label) -> lbl_data
    
    plt_df %>% select(uniprot, protein, condition, LFC_WP) %>% 
      mutate(`PPAT reactivity in D8C` = "Protein expression") %>%
      distinct() -> wp_data
    
    plt_df$protein <- factor(plt_df$protein, levels = targets)
    plt_df %>%
      ggplot(aes(x = condition, y = LFC_tmt_abpp, fill = `PPAT reactivity in D8C`)) +
      geom_hline(yintercept =c(-1, 1),
                 linetype = "dashed",
                 size = LINE_WIDTH) +
      # cysteine data
      geom_point(
        size = POINT_SIZE - 0.75,
        shape = 21,
        color = "black",
        alpha = 0.9,
        stroke = POINT_STROKE
      ) +
      # residue labels
      geom_text(
        data = lbl_data,
        mapping = aes(
          y = LFC_tmt_abpp + case_when(LFC_tmt_abpp > 0 ~ 0.5,
                                       LFC_tmt_abpp < 0 ~ -0.5,
                                       TRUE ~ 0.5),
          color = `PPAT reactivity in D8C`,
          label = residue
        ),
        size = FONT_SIZE,
        size.unit = "pt",
        show.legend = FALSE
      ) +
      # whole proteome data (triangle)
      geom_point(
        data = wp_data,
        aes(x = condition, y = LFC_WP),
        shape = 24,
        color = "black",
        size = POINT_SIZE - 0.75,
        stroke = POINT_STROKE,
        alpha = 0.9
      ) +
      # aesthetics
      my_theme() +
      scale_fill_manual(values = regulation_colors,
                        breaks = names(regulation_colors)) +
      scale_color_manual(values = regulation_colors,
                         breaks = names(regulation_colors)) +
      scale_y_continuous(breaks = round(y_min):round(y_max),
                         limits = c(y_min, y_max)) +
      theme(
        legend.key.spacing.y = unit(-2.5, "mm"),
        legend.margin = margin(b = -15),
        legend.position = "top",
        strip.text = element_text(
          family = FONT_FAMILY,
          color = "black",
          size = FONT_SIZE,
          margin = margin(t = 0)
        ),
        strip.background = element_blank(),
        strip.placement = "outside",
        panel.spacing = unit(0, "mm"),
        legend.title = element_text(margin = margin(b = -2)),
        legend.key = element_blank(),
        legend.title.position = "top",
        legend.key.spacing = unit(0, "mm"),
        legend.spacing.y = unit(-0, "mm"),
        legend.spacing = unit(0, "mm")
      ) +
      labs(y = substitute(
        paste('cysteine reactivity, log'[2], "(FC from D2)"),
        list(n = fil_condition)
      ),
      x = NULL,
      fill = NULL) +
      facet_wrap( ~ protein, ncol = 5, strip.position = "bottom") +
      guides(fill = guide_legend(ncol = 2)) -> plot
    print(plot)
    save_plot(
      paste0(results_dir, "06_dotplots/", title),
      height = fig_height,
      width = fig_width
    )
  }
```

### P53 pathway

```{r, fig.height=2.185, fig.width=2.88}
facetted_multi_condition_dotplot(
  targets = c("DCAF15", "DHCR24","NAP1L1", "NOC2L","FAM111B"),
  display_conditions = c("D4C", "D8C"),
  title = "p53_proteins",
  additional_cysteines_to_label = c("C91", "C511", "C50"),
  fig_height = 2.335 - 0.15,
  fig_width = 2.88
)
```


### PPAT and CDKN1B
```{r, fig.width=1, fig.height=2.2}
facetted_multi_condition_dotplot(
  targets = c("PPAT"),
  display_conditions = c("D8A","D4A","D4C", "D8C"),
  title = "PPAT",
  additional_cysteines_to_label = c("C306"),
  fig_height = 2.2,
  fig_width = 1.4,
  y_min = -2.5,
  y_max = 1.5
)

facetted_multi_condition_dotplot(
  targets = c("CDKN1B"),
  display_conditions = c("D8A", "D8C"),
  title = "CDKN1B",
  additional_cysteines_to_label = c("C29","C148" ),
  fig_height = 2.2,
  fig_width = 1,
  y_min = -1.1, 
  y_max = 1.7
)
```



```{r}
multi_condition_rc_plot <- function(proteins,
                                    ht = 2.5,
                                    wd = 2,
                                    rotate_x_labels = TRUE,
                                    brk_size = 100) {
  paste0(reactivity_vs_wp_dir,
         "/reactivity_wp_comparison_data.csv") %>%
    read.csv() %>%
    filter(protein %in% proteins) %>%
    mutate(residue_loc = 
             str_split_fixed(
               str_split_fixed(residue, "C", 2)[, 2], ";", 2)[, 1] %>% 
             as.numeric()) -> map2k_df
  
  map2k_df %>% filter(tmt_abpp > 200) -> label_data
  
  max_tmt <- max(map2k_df$tmt_abpp)
  map2k_df %>%
    filter(tmt_abpp < 200) %>%
    ggplot(aes(x = reorder(residue, residue_loc), y = tmt_abpp)) +
    theme_minimal() +
    geom_hline(
      yintercept = 100,
      linetype = "solid",
      size = LINE_WIDTH,
      color = "black"
    ) +
    geom_hline(
      yintercept = 200,
      linetype =  "dotted",
      size = LINE_WIDTH,
      color = "black"
    ) +
    ylab("cysteine reactivity (%D2)") +
    xlab("") +
    geom_beeswarm(
      size = POINT_SIZE - 0.5,
      shape = 21,
      col = "black",
      show.legend = FALSE,
      fill = "lightgrey",
      width = 0.05,
      height = 0,
      stroke = POINT_STROKE,
      alpha = 0.7
    ) +
    geom_text(
      data = label_data,
      mapping = aes(
        x = residue,
        y = tmt_abpp + (1.25 / 16) * max_tmt,
        label = condition,
        color = condition
      ),
      inherit.aes = FALSE,
      show.legend = FALSE,
      size = FONT_SIZE,
      size.unit = "pt"
    ) +
    geom_point(
      data = label_data,
      mapping = aes(fill = condition),
      size = POINT_SIZE - 0.5,
      shape = 21,
      col = "black",
      show.legend = FALSE,
      stroke = POINT_STROKE / 2
    ) +
    scale_fill_manual(values = cols) +
    scale_color_manual(values = cols) +
    my_theme() +
    theme(
      strip.background = element_blank(),
      strip.text = element_text(
        color = "black",
        size = 8,
        family = "Arial"
      )
    ) +
    scale_y_continuous(breaks = seq.int(0, 4000, by = brk_size)) +
    facet_wrap(~ protein, scales = "free") -> plt
  if (rotate_x_labels) {
    plt <- plt +
      theme(
        axis.text.x = element_text(
          color = "black",
          angle = 45,
          size = FONT_SIZE,
          family = "Arial",
          vjust = 1,
          hjust = 1
        )
      )
  }
  
  print(plt)
  save_plot(paste0(results_dir, "06_dotplots/", paste(proteins, collapse = "_")),
            height = ht,
            width = wd)
}
```


```{r, fig.width=2, fig.height=1.55}
multi_condition_rc_plot(
  c("HSPA9"),
  ht = 1.55,
  wd = 2,
  rotate_x_labels = FALSE
)

multi_condition_rc_plot(
  c("LONP1"),
  ht = 1.55,
  wd = 2,
  rotate_x_labels = FALSE
)


```

```{r, fig.width=1.3, fig.height=2}
multi_condition_rc_plot(
  c("PIP5K1B"),
  ht = 2,
  wd = 1.3,
  rotate_x_labels = FALSE,
  brk_size = 200
)

```


```{r, fig.width = 2.5, fig.height=2.5}

multi_condition_rc_plot(
  proteins <- c("MAP2K3", "MAP2K4"),
  ht = 2.5,
  wd = 2.5,
  rotate_x_labels = TRUE,
  brk_size = 100
)
```




## Quantifed cysteines barplot

```{r, fig.height=1.8, fig.width=2}

AXIS_TXT_MRGN = -5


quant_colors <- c(
  "reactivity changes" = "indianred2",
  "analyzed" = "grey39",
  "not analyzed" = "grey90"
)

"/Users/henrysanford/dev/reactivity/05_reactivity_changes/output/rc_df.csv" %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(rc_any = case_when(((num_reactivity_changes_D4C != 0) |
                               (num_reactivity_changes_D8C != 0) |
                               (num_reactivity_changes_D4A != 0) |
                               (num_reactivity_changes_D8A != 0)
  ) ~ TRUE,
  TRUE ~ FALSE)) %>% select(c(
    uniprot,
    rc_any,
    num_quantified_peptides,
    num_reactivity_changes_D4C
  )) -> df
print(df)
df %>%
  mutate(
    reactivity_change = case_when(
      num_quantified_peptides == 1 ~ "not analyzed",
      rc_any == TRUE ~ "reactivity changes",
      TRUE ~ "analyzed"
    )
  ) %>%
  mutate(binned_quantified_peptides = case_when(
    (num_quantified_peptides > 6) ~ "7+",
    (num_quantified_peptides > 4) ~ "5-6",
    TRUE ~ as.character(num_quantified_peptides)
  )) -> df
print(df)
df$reactivity_change <-
  factor(df$reactivity_change, levels = rev(names(quant_colors)))

df %>%
  arrange(reactivity_change, ascending = TRUE) %>%
  distinct(.keep_all = TRUE) %>%
  ggplot(aes(x = binned_quantified_peptides, fill = reactivity_change)) +
  geom_bar(show.legend = FALSE,
           width = BARPLOT_WIDTH) +
  scale_fill_manual(values = quant_colors,
                    breaks = names(quant_colors),
                    name = "") +
  guides(fill = guide_legend(ncol = 1, hjust = 1)) +
  scale_y_break(c(100, 300), scales = 1 / 1.2, expand = c(0, 0)) +
  scale_x_discrete() +
  my_theme() +
  theme(
    legend.position =  "top",
    legend.justification = "left",
    legend.key.spacing = unit(-0.1, "mm"),
    legend.key.spacing.x = unit(500, "mm"),
    legend.margin = margin(b = -5),
    # reverse annoying ggbreak changes
    axis.line.y.right = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title.x = element_text(
      margin = margin(t = AXIS_TXT_MRGN),
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE,
      hjust = 0.5
    ),
    axis.title.y = element_text(
      margin = margin(r = AXIS_TXT_MRGN),
      color = "black",
      family = FONT_FAMILY,
      size = FONT_SIZE,
      vjust = 0.5
    ),
    panel.border = element_blank(),
    axis.line.x = AXIS_LINE,
    axis.line.y = AXIS_LINE
  ) +
  ylab("number of proteins") +
  xlab("quantified cysteines") +
  ylim(0, 2400)

save_plot(
  paste0(results_dir, "07_barplots/num_peptides_rc"),
  height = 2,
  width = 2
)

```

## Reactivity change stacked barplots

```{r, fig.width=2.7, fig.height=2}

colors <- c(
  "Other" = "grey90",
  "Nucleotide binding   \nproteins" = "#96D1A9"
)

x_axis_exand <- 0.01

directions <- c("Higher", "Lower")
paste0(protein_list_dir, "GO_term_lists/Mitochondrial/mitochondrial_gene_list_go.csv") %>%
  read_csv() %>% 
  pull(uniprot) -> mito_uniprots

directions <- c("Higher", "Lower")
paste0(protein_list_dir, "GO_term_lists/Nucleotide binding/02_genes_matching_query.csv") %>%
  read_csv() %>% 
  pull(uniprot) -> nuc_uniprots

paste0(reactivity_vs_wp_dir, "/cysteine_reactivity_changes.csv") %>%
  read_csv() -> cys_df
cys_df$direction_of_reactivity_change <- factor(cys_df$direction_of_reactivity_change,directions)

cys_df %>%
  mutate(mitochondrial = case_when(uniprot %in% nuc_uniprots ~ "Nucleotide binding   \nproteins",
                                   TRUE ~ "Other")) -> cys_df
cys_df$mitochondrial <- factor(cys_df$mitochondrial, levels = names(colors))
cys_df$condition <- factor(cys_df$condition, levels = c("D4A", "D8A", "D4C", "D8C"))
cys_df %>%
  ggplot(aes(x = condition, fill = mitochondrial)) -> mito_plot
mito_plot +
    geom_bar(
           stat="count", show.legend = TRUE, width = 0.7, col = "black", size = LINE_WIDTH) + 
  geom_text(
    data = cys_df %>% filter(direction_of_reactivity_change == "Higher"),
    stat = "count", 
            inherit.aes = FALSE,
            mapping = aes(x = condition, label = ..count..), 
            vjust = -0.1, 
            size = FONT_SIZE, 
            size.unit = "pt") + 
    geom_text(
    data = cys_df %>% filter(direction_of_reactivity_change == "Lower"),
    stat = "count", 
            inherit.aes = FALSE,
            mapping = aes(x = condition, label = ..count..), 
            vjust = 1.1, 
            size = FONT_SIZE, 
            size.unit = "pt") + 
  facet_grid("direction_of_reactivity_change", scales = "free_y", space = "free") + 
  guides(fill = guide_legend(ncol =1)) +  
  legend_theme() +
  my_theme() +
  theme(
    strip.background = element_blank(),
    panel.spacing = unit(0, "mm"),
  ) +
  xlab("") +
  scale_fill_viridis_d(direction = -1) +
  scale_fill_discrete(type = colors) +
  ylab("number of reactivity changes") + 
  scale_y_continuous(expand = c(0,0)) +
  facetted_pos_scales(
    y = list(
      scale_y_continuous(expand = expansion(mult = c(0,x_axis_exand)), limits = c(0, 79)),
      scale_y_continuous(expand = expansion(mult = c(x_axis_exand,0)), trans = "reverse", limits = c(90,0))
      )
    )
save_plot(
  fn = paste0(results_dir, "07_barplots/nuc_binding/nuc_binding_split"),
  width = 2.7,
  height = 2
)
```


```{r, fig.width=2.5, fig.height=1.5}

colors <- c(
  "Other" = "grey90",
  "Mitochondrial" = "#21918c"
)

directions <- c("Higher", "Lower")
paste0(protein_list_dir, "GO_term_lists/Mitochondrial/mitochondrial_gene_list_go.csv") %>%
  read_csv() %>% 
  pull(uniprot) -> mito_uniprots

paste0(reactivity_vs_wp_dir, "/cysteine_reactivity_changes.csv") %>%
  read_csv() -> cys_df
cys_df$direction_of_reactivity_change <- factor(cys_df$direction_of_reactivity_change,directions)

cys_df %>%
  mutate(mitochondrial = case_when(uniprot %in% mito_uniprots ~ "Mitochondrial",
                                   TRUE ~ "Other")) -> cys_df
cys_df$mitochondrial <- factor(cys_df$mitochondrial, levels = names(colors))
cys_df$condition <- factor(cys_df$condition, levels = c("D4A", "D8A", "D4C", "D8C"))
cys_df %>%
  ggplot(aes(x = condition, fill = mitochondrial)) -> mito_plot
mito_plot +
    geom_bar(
           stat="count", show.legend = TRUE, width = 0.7, col = "black", size = LINE_WIDTH) + 
  geom_text(
    data = cys_df %>% filter(direction_of_reactivity_change == "Higher"),
    stat = "count", 
            inherit.aes = FALSE,
            mapping = aes(x = condition, label = ..count..), 
            vjust = -0.1, 
            size = FONT_SIZE, 
            size.unit = "pt") + 
    geom_text(
    data = cys_df %>% filter(direction_of_reactivity_change == "Lower"),
    stat = "count", 
            inherit.aes = FALSE,
            mapping = aes(x = condition, label = ..count..), 
            vjust = 1.1, 
            size = FONT_SIZE, 
            size.unit = "pt") + 
  facet_grid("direction_of_reactivity_change", scales = "free_y", space = "free") + 
  guides(fill = guide_legend(ncol =1)) +  
  legend_theme() +
  my_theme() +
  theme(
    strip.background = element_blank(),
    panel.spacing = unit(0, "mm"),
  ) +
  xlab("") +
  scale_fill_viridis_d(direction = -1) +
  scale_fill_discrete(type = colors) +
  ylab("number of reactivity changes") + 
  scale_y_continuous(expand = c(0,0)) +
  facetted_pos_scales(
    y = list(
      scale_y_continuous(expand = expansion(mult = c(0,x_axis_exand)), limits = c(0, 88), breaks = c(0, 25, 50, 75)),
      scale_y_continuous(expand = expansion(mult = c(x_axis_exand,0)), trans = "reverse", limits = c(101,0))
      )
    )
save_plot(
  fn = paste0(results_dir, "07_barplots/mitochondrial/mitochondrial_split"),
  width = 2.5,
  height = 1.5
)
```

## Expression and reactivity boxplot

```{r}
expression_and_reactivity_boxplot <- function(condition_to_display,
                                              control_condition,
                                              signal_intensity_col_prefix,
                                              proteins_to_plot,
                                              whole_proteome_table,
                                              reactivity_table) {
  whole_proteome_table %>%
    # select columns needed for plot
    select(starts_with(signal_intensity_col_prefix) | "protein") %>%
    select(-c("description")) %>%
    filter(protein %in% proteins_to_plot) %>%
    # concat protein and residue
    column_to_rownames("protein") -> si_matrix
  
  # calculate fold change to median of control condition
  fold_change = si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =
                                    T)
  fold_change %>%
    rownames_to_column("protein") %>%
    # pivot to long format for plotting
    pivot_longer(cols = colnames(.)[-1],
                 names_to = "condition_replicate",
                 values_to = "channel_ratio") %>%
    # drop missing values
    drop_na() %>%
    separate_wider_delim(
      cols = c("condition_replicate"),
      delim = "_",
      names = c("condition", "technical_replicate", "experiment"),
      too_many = "drop"
    ) %>%
    filter(protein %in% proteins_to_plot) %>%
    filter(condition == condition_to_display) -> wp_df
  
  wp_df$median_rc <- 0
  wp_df$residue <- "Protein"
  wp_df$regulation <- "Protein expression"
  wp_df$residue_loc <- 0
  
  wp_df %>%
    group_by(protein) %>%
    mutate(median_wp = median(channel_ratio)) -> wp_df
  wp_df %>% select(protein, median_wp) -> wp_stats
  
  reactivity_table %>%
    # select columns needed for plot
    select(starts_with(signal_intensity_col_prefix) |
             "protein" | "residue") %>%
    select(-c("description")) -> si_matrix
  si_matrix$residue <- gsub(';', ',', si_matrix$residue)
  si_matrix %>%
    # concat protein and residue
    unite(rowname, protein, residue) %>%
    column_to_rownames() -> si_matrix
  # calculate fold change to median of AB condition
  fold_change = si_matrix / apply(si_matrix[, grepl(control_condition, colnames(si_matrix))], 1, median, na.rm =
                                    T)
  fold_change %>%
    rownames_to_column("protein_residue") %>%
    # pivot to long format for plotting
    pivot_longer(cols = colnames(.)[-1],
                 names_to = "condition_replicate",
                 values_to = "channel_ratio") %>%
    # drop missing values
    drop_na() %>%
    separate_wider_delim(
      cols = c("protein_residue"),
      delim = "_",
      names = c("protein", "residue")
    ) %>%
    separate_wider_delim(
      cols = c("condition_replicate"),
      delim = "_",
      names = c("condition", "experiment", "technical_replicate"),
      too_many = "drop"
    ) %>%
    filter(protein %in% proteins_to_plot) %>%
    filter(condition == condition_to_display) -> df
  
  df$protein <- factor(df$protein, levels = proteins_to_plot)
  df$residue_loc <- str_split_fixed(df$residue, "C", 2)[, 2]
  df$residue_loc <-
    str_split_fixed(df$residue_loc, ",", 2)[, 1] %>% as.numeric()
  df %>%
    group_by(protein, residue) %>%
    mutate(median_rc = median(channel_ratio)) -> rc_df
  
  
  paste0(reactivity_vs_wp_dir, "/rc_df_long_format.csv") %>% read_csv() -> full_df
  
  rc_df %>%
    merge(wp_stats, by = "protein", all.x = TRUE) %>%
    mutate(wp_rc_ratio =  (median_rc) / (median_wp),) -> rc_df
  merge(
    x = rc_df,
    y = full_df %>%
      filter(condition == condition_to_display) %>%
      select(protein, residue, reactivity_change, LFC_tmt_abpp),
    by = c("protein", "residue"),
    all.x = TRUE
  ) %>%
    mutate(
      regulation = case_when(
        if_all(reactivity_change, is.na) ~ "Unchanged",
        LFC_tmt_abpp > 0 ~ "Higher",
        LFC_tmt_abpp < 0 ~ "Lower",
        TRUE ~ "Unchanged"
      )
    ) -> plt_df
  
  wp_df$reactivity_change = FALSE
  wp_df$LFC_tmt_abpp = 0
  wp_df$wp_rc_ratio = 0
  wp_df$regulation = "Protein expression"
  
  rbind(plt_df, wp_df) -> df
  df$protein <- factor(df$protein, levels = proteins_to_plot)
  df %>%
    unique() %>%
    ggplot(aes(
      x = reorder(residue, residue_loc),
      fill = regulation,
      y = log2(channel_ratio),
      shape = regulation
    )) +
    geom_hline(
      yintercept = 0,
      size = LINE_WIDTH,
      linetype = "solid",
      color = "grey90"
    ) +
    geom_boxplot(
      outliers = FALSE,
      width = 0.7,
      lwd = LINE_WIDTH,
      show.legend = FALSE,
      alpha = 0.8
    ) +
    geom_beeswarm(
      corral.width = 1.2,
      cex = 2,
      col = "black",
      stroke = POINT_STROKE,
      show.legend = FALSE,
      size = 1,
      alpha = 0.9
    ) +
    facet_rep_grid( ~ protein, scales = "free_x", space = "free",) +
    my_theme() +
    theme(
      plot.title = element_text(
        size = 6,
        color = "black",
        family = "Helvetica",
        hjust = 0.5
      ),
      strip.background = element_blank(),
      axis.text.x = element_text(
        color = "black",
        angle = 45,
        size = FONT_SIZE,
        family = "Helvetica",
        vjust = 1,
        hjust = 1
      ),
      panel.spacing = unit(2, "mm"),
      panel.border = element_blank(),
      strip.text = element_text(
        size = 8,
        family = "Helvetica",
        color = "black"
      ),
      axis.line.x = AXIS_LINE,
      axis.line.y = AXIS_LINE,
      strip.clip = "off"
    ) +
    labs(y = substitute(
      paste('log'[2],
            "(fold change):", n, "/", m),
      list(n = condition_to_display, m = control_condition)
    ),
    x = "",
    fill = "Reactivity") +
    scale_fill_manual(
      values = c(
        "Higher" = SIG_UP_COL,
        "Lower" = SIG_DOWN_COL,
        "Unchanged" = UNCHANGED_COL,
        "Protein expression" = "#54A868"
        
      )
    ) +
    scale_shape_manual(values = c(
      "Higher" = 21,
      "Unchanged" = 21,
      "Lower" = 21,
      "Protein expression" = 24
      
    )) -> plot
  ht <- 2.2
  wd <- 8
  file_name <- paste(
    condition_to_display,
    "vs",
    control_condition,
    paste(proteins_to_plot, collapse = "_"),
    sep = "_"
  )
  
  print(plot)
  
  save_plot(
    paste0(results_dir, "05_boxplots/", file_name),
    width = wd,
    height = ht
  )
  return(plt_df)
}
```

```{r, fig.width=8, fig.height=2}
'../../data/whole_proteome/combined_files/07_combfiles_percctrl_passes_two_replicate_variability_filter.csv' %>%
  read_csv(show_col_types = FALSE) -> wp_table

'../../data/reactivity/combined_files/20240821_5_rep_variability_filter/07_combfiles_cysaggr_percctrl_passes_two_replicate_variability_filter.csv' %>%
  read_csv(show_col_types = FALSE) -> rc_table

target_proteins <- c(
  "NDUFV1",
  "NDUFB7"
) 
condition_to_display <- "D4C"
control_condition <- "D2"
for(cond in conditions){
  signal_intensity_col_prefix <- "D"
  df <- expression_and_reactivity_boxplot(
  condition_to_display = cond,
  control_condition = control_condition,
  signal_intensity_col_prefix = "D",
  whole_proteome_table = wp_table,
  reactivity_table = rc_table,
  proteins_to_plot = target_proteins)
}
```


## Add back experiment

Aesthetic parameters for the add back experiment

```{r}
add_back_fills <- c(
  "D2" = "darkgrey",
  "D2-ATP" = "grey99",
  "spacer1" = "transparent",
  "D8A" = "#603EA6",
  "D8A-ATP" = "#D13B95",
  "spacer2" = "transparent",
  "D8C" = "#E27753",
  "D8C-ATP" = "#EF3D3B"
)

add_back_cols <- c(
  "D2" = "black",
  "D2-ATP" = "black",
  "spacer1" = "transparent",
  "D8A" = "black",
  "D8A-ATP" = "black",
  "spacer2" = "transparent",
  "D8C" = "black",
  "D8C-ATP" = "black"
)

add_back_shapes <- c(
  "D2" = 21,
  "D2-ATP" = 23,
  "spacer1" = 21,
  "D8A" = 21,
  "D8A-ATP" = 23,
  "spacer2" = 21,
  "D8C" = 21,
  "D8C-ATP" = 23
)
```


### boxplot

```{r, fig.height=2.2, fig.width=6.5}
facetted_add_back_boxplot <-
  function(proteins_to_plot,
           fig_width,
           fig_height,
           show_legend = TRUE) {
    paste0(add_back_dir, 'boxplots/formatted_reactivity_data.csv') %>%
      read_csv(show_col_types = FALSE) %>%
      filter(protein %in% proteins_to_plot) %>%
      mutate(
        lfc = log2(percent_control / 100),
        super_condition <-
          str_split_fixed(condition, "-ATP", 2)[, 1]
      ) %>%
      select(protein,
             residue,
             residue_loc,
             lfc,
             condition) -> df
    
    # add invisible spacers separating conditions D2, D8A, D8C
    for (plot_protein in proteins_to_plot) {
      map <- setNames(unique(df %>% filter(protein == plot_protein) %>% pull(residue)),
                      unique(df %>% filter(protein == plot_protein) %>% pull(residue_loc)))
      for (residue_location in names(map)) {
        for (spacer_num in c("spacer1", "spacer2")) {
          df[nrow(df) + 1, ] = list(
            "protein" = plot_protein,
            "residue" = map[residue_location],
            "residue_loc" = as.double(residue_location),
            "lfc" = 0,
            "condition" = spacer_num
          )
        }
      }
    }
    
    legend_breaks = names(add_back_fills)[!grepl("spacer", names(add_back_fills))]
    
    df %>%
      mutate(condition = factor(df$condition, levels = names(add_back_cols))) %>%
      arrange(condition) %>%
      ggplot(aes(
        x = reorder(residue, residue_loc),
        y = lfc,
        fill = condition,
        color = condition,
        shape = condition
      )) +
      geom_hline(yintercept = c(1, -1),
                 linetype = "dashed",
                 size = LINE_WIDTH) +
      geom_boxplot(
        alpha = 0.9,
        lwd = LINE_WIDTH,
        outliers = FALSE,
        show.legend = FALSE
      ) +
      geom_jitter(
        size = POINT_SIZE - 1.5,
        alpha = 0.9,
        stroke = POINT_STROKE,
        show.legend = show_legend,
        position = position_jitterdodge(jitter.width = 0.1)
      ) +
      labs(y = substitute(paste(
        'cysteine reactivity, log'[2],
        "(FC from D2)"
      )),
      x = "") +
      guides(fill = element_blank()) +
      my_theme() +
      theme(
        strip.background = element_blank(),
        strip.text = element_text(
          color = "black",
          size = FONT_SIZE,
          family = "Arial"
        ),
        plot.title = element_text(
          size = 8,
          color = "black",
          family = "Arial",
          hjust = 0.5
        ),
        legend.title = element_blank(),
        legend.box.background = element_blank(),
        legend.position = "top",
        legend.key = element_blank(),
        legend.margin = margin(b = -14, t = -10),
        legend.key.spacing.y = unit(-3, "mm"),
        legend.text = element_text(
          size = 6,
          color = "black",
          family = "Arial",
          margin = margin(l = -4)
        ),
        legend.key.spacing.x = unit(-1, "mm")
      ) +
      scale_fill_manual(values = add_back_fills, breaks = legend_breaks) +
      scale_color_manual(values = add_back_cols, breaks = legend_breaks) +
      scale_shape_manual(values = add_back_shapes, breaks = legend_breaks) +
      geom_vline(
        xintercept = 1.5:length(unique(df$residue)),
        size = LINE_WIDTH,
        linetype = "dotted"
      ) +
      #scale_y_continuous(limits =c(-3.2,3.2), breaks = c(-3,-2,-1,0,1,2,3) ) +
      facet_grid( ~ protein, space = "free", scales = "free_x") -> plt
    print(plt)
    
    save_plot(paste0(
      add_back_dir,
      'boxplots/',
      paste(proteins_to_plot, collapse = "_")
    ),
    width = fig_width,
    height = fig_height)
  }
```



```{r, fig.width=2.5, fig.height=2.2}
add_back_targs <- c("HSPA9", "LONP1","MAP2K4")
for (target in add_back_targs){
 facetted_add_back_boxplot(
   proteins_to_plot = c(target),
   fig_width = 2.5,
   fig_height = 2.2
 )
}
```

```{r, fig.width=3.5, fig.height=2.7}
facetted_add_back_boxplot(
 proteins_to_plot = c("MAP2K4", "MAP2K3"),
 fig_width = 3.5,
 fig_height = 2.4,
 show_legend = TRUE
)
```


```{r, fig.width=3.5, fig.height=2.4}
facetted_add_back_boxplot(
  proteins_to_plot = c("ZAP70"),
  fig_width = 3.5,
  fig_height = 2.4
)
```


### condensed ATP sensitivity barplot

```{r, fig.height=1.8, fig.width=2.5}
condensed_atp_barplot_direction <- function(proteins_to_plot, fig_width, fig_height, rotate_x_labels = TRUE){
  dodge_width <- 0.8
  paste0(add_back_dir, 'protein_table_ATP_add_back_2_reps.csv') %>% 
      read_csv(show_col_types = FALSE) %>%
      filter(protein %in% proteins_to_plot) %>%
      pivot_longer(
        cols = c("ratio_D2_+/-_ATP","ratio_D8A_+/-_ATP","ratio_D8C_+/-_ATP"), 
        names_to = "super_condition", values_to = "ratio") %>%
      mutate(
        residue_loc = as.numeric(
          str_split_fixed(
            str_split_fixed(residue, "C", n = 2)[,2]
            , ",",n = 3)[,1]),
        super_condition = str_split_fixed(super_condition, "_",n = 3)[,2],
        general_reactivity_change = 
        case_when(
          (super_condition == "D8A") & (log2(`median_D8A-MgCl2` / 100) > 1) ~ "Higher",
          (super_condition == "D8A") & (log2(`median_D8A-MgCl2` / 100) < -1) ~ "Lower",
          (super_condition == "D8C") & (log2(`median_D8C-MgCl2` / 100) > 1) ~ "Higher",
          (super_condition == "D8C") & (log2(`median_D8C-MgCl2` / 100) < -1) ~ "Lower",
          TRUE ~ "Unchanged"
          ),
        super_condition = factor(super_condition, levels = c("D2", "D8A", "D8C")),
        protein = factor(protein, levels = proteins_to_plot)) %>%
    ggplot(aes(x = reorder(residue, residue_loc, group = super_condition), y = log2(ratio))) +
    scale_fill_manual(values = EXHAUSTION_COLS) + 
    geom_bar(
      mapping = aes(fill = super_condition),
      stat = "identity", 
      position = position_dodge(width = dodge_width), 
      width = 0.7, 
      color = "black", size = LINE_WIDTH) + 
    ggnewscale::new_scale_fill() +
    scale_fill_manual(values = c("Higher" = SIG_UP_COL, "Lower" = SIG_DOWN_COL)) + 
    # symbols indicating change from D2
    geom_point(
      mapping = aes(
        fill = general_reactivity_change,
        color = general_reactivity_change,
        group = super_condition,
      alpha = general_reactivity_change,
      shape = general_reactivity_change,
       y = log2(ratio) + ((log2(ratio) * 1/abs(log2(ratio)))*0.1),
      ),
      position = position_dodge2(width = dodge_width),
      show.legend = FALSE,
      stroke = POINT_STROKE,
      size = 1.2,
      color = "black"
    ) +
    facet_grid(~protein, scales = "free_x", space = "free") + 
    my_theme() + 
    scale_alpha_manual(values = c(
      "Unchanged" = 0,
      "Higher" = 1,
      "Lower" = 1
    )) +
    scale_shape_manual(values = c(
      "Unchanged" = 0,
      "Higher" = 24,
      "Lower" = 25
    )) +
    labs(
      x = NULL,
      y = expression("cysteine reactivity, log"[2]*"(+/- ATP)")
    ) + 
      theme(
        strip.background = element_blank(),
        strip.text = element_text(color="black", size=FONT_SIZE, family="Arial"),
        plot.title = element_text(size=8, color = "black", family="Arial", hjust=0.5),
        legend.title=element_blank(),
        legend.box.background = element_blank(),
        legend.position = "top",
        legend.key = element_blank(),
        legend.margin = margin(b = -10),
        strip.clip = "off",
        legend.key.width  = unit(2, "mm"),
        legend.key.height  = unit(2, "mm"),
        legend.text = element_text(size = 6, color = "black", family = "Arial", margin = margin(l = 0)),
        legend.key.spacing.x = unit(0, "mm"),
        axis.text.x = element_text(size = 6, color = "black", family = "Arial")
      ) + 
    geom_hline(yintercept = 0, color="grey", size = LINE_WIDTH) + 
    geom_hline(yintercept = -1, color="black", size = LINE_WIDTH, linetype = "dashed") -> plot
  if(rotate_x_labels){
    plot + theme(axis.text.x = element_text(
          color="black",angle=45, size=FONT_SIZE, family="Arial", vjust = 1, hjust=1)) -> plot
  }
  print(plot)
  save_plot(
    paste0(add_back_dir, "/", paste(proteins_to_plot, collapse = "_")),
    width = fig_width, height = fig_height
  )
}

```

```{r, fig.width=2.5, fig.height=2.055}
condensed_atp_barplot_direction(
  c(
    "AFG3L2", "SPG7", "HPRT1" 
  ),
  fig_width = 2.5,
  fig_height = 2.055,
  rotate_x_labels = FALSE
)

```

```{r, fig.width=3.2, fig.height=2.4}
condensed_atp_barplot_direction(
  c("GCN1"),
  fig_width = 3.2,
  fig_height = 2.4,
  rotate_x_labels = TRUE
)
```



### stacked up and down barplot

```{r, fig.width = 1.4, fig.height=1.7}
x_axis_expand <- 0.2

nucleotide_cols <- c("Nucleotide binding" = "#96d1a9",
                     "Other" = UNCHANGED_COL)

paste0(add_back_dir, 'bargraphs/formatted_add_back_regulation.csv') %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(nucleotide_binding = case_when(uniprot %in% nuc_uniprots ~ "Nucleotide binding",
                                        TRUE ~ "Other")) %>%
  filter(value != "Unchanged") -> df
df$nucleotide_binding <-
  factor(df$nucleotide_binding, levels = rev(names(nucleotide_cols)))
df$condition <-
  gsub(pattern = "_\\+\\/\\-_ATP",
       replacement = "",
       x = 	df$condition)

df %>%
  arrange(nucleotide_binding) %>%
  ggplot(aes(x = condition, fill = nucleotide_binding)) +
  geom_bar(
    stat = "count",
    width = 0.7,
    color = "black",
    size = LINE_WIDTH,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = nucleotide_cols) +
  geom_text(
    data = df %>% filter(value == "Higher"),
    stat = "count",
    aes(label = ..count..,
        y = ..count..,
        x = condition),
    vjust = -.5,
    inherit.aes = FALSE,
    size = FONT_SIZE,
    size.unit = "pt",
    position = position_dodge(width = 0.9)
  ) +
  geom_text(
    data = df %>% filter(value == "Lower"),
    stat = "count",
    aes(
      label = after_stat(count),
      y = after_stat(count),
      x = condition
    ),
    vjust = 1.1,
    inherit.aes = FALSE,
    size = FONT_SIZE,
    size.unit = "pt",
    position = position_dodge(width = 0.9)
  ) +
  legend_theme() +
  my_theme() +
  theme(
    strip.background = element_blank(),
    axis.text.x = element_text(
      size = 6,
      color = "black",
      family = "Arial"
    )
  ) +
  facet_grid("value", scales = "free_y", space = "free") +
  ylab("number of cysteines") +
  xlab("") +
  theme(strip.background = element_blank(),
        panel.spacing = unit(0, "mm"),) +
  facetted_pos_scales(y = list(
    scale_y_continuous(expand = expansion(mult = c(0, x_axis_expand)), limits = c(0, 160)),
    scale_y_continuous(
      expand = expansion(mult = c(x_axis_expand, 0)),
      trans = "reverse",
      limits = c(150, 0)
    )
  ))

save_plot(
  paste0(add_back_dir, "bargraphs/sites_affected_by_add_back_split"),
  height = 2,
  width = 1.4
  
)
```

### ATP sensitivity scatterplot

```{r, fig.width=2, fig.height=3}
atp_sensitivity_scatterplot <- function(condition,
                                        control,
                                        y_max,
                                        y_min,
                                        x_max,
                                        x_min,
                                        fill_variable = "ATP_ratio") {
  y_nudge <- 3
  teal <- paste0("More ATP sensitive at ", condition)
  yellow <- paste0("More ATP sensitive at ", control)
  
  
  line_df <- data.frame(
    x1 = c(y_min - 1 , y_min + 1),
    x2 = c(x_max, x_max),
    y1 = c(y_min, y_min),
    y2 = c(x_max + 1, x_max - 1)
  )
  
  
  paste0(add_back_dir, 'protein_table_ATP_add_back_2_reps.csv') %>%
    read_csv(show_col_types = FALSE) %>%
    mutate(name = paste0(protein, "_", residue)) -> df
  df$x_var <- log2(df[, paste0("ratio_", condition, "_+/-_ATP")][[1]])
  df$y_var <- log2(df[, paste0("ratio_", control, "_+/-_ATP")][[1]])
  # drop Inf
  df <- df[(is.finite(df$y_var) & is.finite(df$x_var)), ]
  
  df$quadrant <-
    df[, paste0("ATP_sensitivity_", condition, "_vs_", control)][[1]]
  if (fill_variable == "ATP_ratio") {
    n_cols <- 1
    df$regulation <- df$quadrant
    atp_ratio_cols <- setNames(c("#21918c", "#bade28", "grey"),
                               c(teal, yellow, "Unchanged"))
    df %>% filter(
      name %in% c(
        "AFG3L2_C313",
        "HPRT1_C23",
        "C106",
        "HSPA9_C317",
        "LONP1_C682",
        "MAP2K3_C305",
        "MAP2K4_C246",
        "CHUK_C30"
      )
    ) -> lbl_df
    
  } else {
    n_cols <- 4
    df$regulation <- df$proteasome_subunit_family
    atp_ratio_cols <- c(
      "PSMA" = "#bade28",
      "PSMB" = "#21918c",
      "PSMC" = "#440154",
      "PSMD" = "#fc8961",
      "Other" = "gray88"
    )
    df$regulation <- factor(df$regulation, rev(names(atp_ratio_cols)))
    df %>% filter(regulation != "Other") -> lbl_df
  }
  
  print(paste(
    max(df$y_var[is.finite(df$y_var)]),
    min(df$y_var[is.finite(df$y_var)]),
    max(df$x_var[is.finite(df$x_var)]),
    min(df$x_var[is.finite(df$x_var)])
  ))
  
  df %>%
    arrange(regulation) %>%
    ggplot(aes(x = x_var,
               y = y_var)) +
    geom_point(
      aes(fill = regulation),
      shape = 21,
      size = POINT_SIZE - 1,
      stroke = POINT_STROKE / 2,
      color = "gray88",
      alpha = 0.4
    ) +
    guides(
      fill = guide_legend(ncol = n_cols, reverse = TRUE),
      color = NULL,
      alpha = NULL,
      size = NULL
    ) +
    my_theme() +
    theme(
      legend.title = element_blank(),
      legend.box.background = element_blank(),
      legend.position = "top",
      legend.justification.top = "top",
      legend.justification = "center",
      legend.box.just = "center",
      aspect.ratio = 1,
      legend.key = element_blank(),
      legend.margin = margin(l = -5, b = -5.5),
      legend.box.margin = margin(l = -5, b = -0.8),
      legend.key.spacing.y = unit(-0.9, "cm"),
      legend.key.spacing.x = unit(-0.05, "cm")
    ) +
    geom_segment(
      data = line_df,
      aes(
        x = x1,
        y = y1,
        xend = x2,
        yend = y2
      ),
      inherit.aes = FALSE,
      linetype = "dashed",
      size = LINE_WIDTH,
      show.legend = FALSE,
      color = "black"
    ) +
    geom_text_repel(
      data = lbl_df %>% filter(quadrant == teal),
      mapping = aes(
        x = x_var,
        y = y_var,
        label = name,
        color = regulation
      ),
      size = FONT_SIZE_MM,
      nudge_y = y_nudge,
      segment.size = LINE_WIDTH,
      show.legend = FALSE
    ) +
    geom_text_repel(
      data = lbl_df %>% filter(quadrant == yellow),
      mapping = aes(
        x = x_var,
        y = y_var,
        label = name,
        color = regulation
      ),
      size = FONT_SIZE_MM,
      nudge_y = -y_nudge,
      segment.size = LINE_WIDTH,
      show.legend = FALSE
    ) +
    scale_fill_manual(values = atp_ratio_cols) +
    scale_color_manual(values = atp_ratio_cols) +
    labs(x = substitute(
      paste('cysteine reactivity, log'[2], "(", n, "-ATP/", n, ")"),
      list(n = condition)
    ),
    y = substitute(
      paste('cysteine reactivity, log'[2], "(", n, "-ATP/", n, ")"),
      list(n = control)
    ),) +
    scale_x_continuous(
      breaks = round(x_min):round(x_max),
      limits = c(x_min, x_max),
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      breaks = round(y_min):round(y_max),
      limits = c(y_min, y_max),
      expand = c(0, 0)
    ) -> plot
  
  print(plot)
  
  save_plot(
    paste0(
      add_back_dir,
      "scatterplot/add_back_scatterplot_",
      condition,
      "_vs_",
      control
    ),
    width = 2,
    height = 3
  )
}

for (fill_var in c("ATP_ratio", "PSM_family")) {
  atp_sensitivity_scatterplot(
    condition = "D8C",
    control = "D8A",
    y_max = 4.9,
    y_min = -3.5,
    x_max = 3.4,
    x_min = -6,
    fill_variable = fill_var
  )
  
  atp_sensitivity_scatterplot(
    condition = "D8C",
    control = "D2",
    y_max = 4.4,
    y_min = -2.9,
    x_max = 3.4,
    x_min = -6,
    fill_variable = fill_var
  )
  
  atp_sensitivity_scatterplot(
    condition = "D8A",
    control = "D2",
    y_max = 6.5,
    y_min = -3.5,
    x_max = 4.8,
    x_min = -6,
    fill_variable = fill_var
  )
}
```


