# Helper methods
```{r}
# set working directory to script location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# load figure settings
source("../../bin/figure_parameters.R")

rc_dotplot <- function(df,targets,
                       title,
                       ratio_aspect = 1,
                       ht = 4,
                       wd = 2,
                       y_max = 3.2,
                       y_min = -3.2, 
                       control_condition = "D2", 
                       plot_dir = "/Users/henrysanford/dev/test_data/later_timepoints_reactivity/processed_results/dotplots/") {
  plt_df <- df %>%
    filter(condition == fil_condition) %>%
    filter(protein %in% targets) %>%
    mutate(
      `Cysteine reactivity in D8C` = case_when(
        reactivity_change != "true" ~ "Unchanged",
        LFC_tmt_abpp > 0 ~ "Higher",
        LFC_tmt_abpp < 0 ~ "Lower",
        TRUE ~ "Unchanged"
      ),
      protein = factor(protein, levels = targets)
    )
  
  lbl_data <- plt_df %>%
    filter(reactivity_change == "true")
  
  wp_data <- plt_df %>% 
    select(uniprot, protein, condition, LFC_WP) %>%
    mutate(`Cysteine reactivity in D8C` = "Protein expression") %>% 
    distinct()
  
  print(max(abs(plt_df$LFC_tmt_abpp)))
  print(max(abs(plt_df$LFC_WP)))

  plt_df %>%
    arrange(by = "protein") %>%
    ggplot(aes(x = protein, y = LFC_tmt_abpp, fill = `Cysteine reactivity in D8C`)) +
    geom_hline(yintercept = c(1,-1),
               linetype = "dashed",
               size = LINE_WIDTH) +
    # cysteine data
    geom_point(
      size = POINT_SIZE - 0.75,
      shape = 21,
      color = "black",
      alpha = 0.9,
      stroke = POINT_STROKE
    ) +
    # residue labels
    geom_text(
      data = lbl_data,
      mapping = aes(
        y = LFC_tmt_abpp + case_when(
          LFC_tmt_abpp > 0 ~ lbl_nudge,
          LFC_tmt_abpp < 0 ~ -lbl_nudge,
          TRUE ~ lbl_nudge
        ),
        color = `Cysteine reactivity in D8C`,
        label = residue
      ),
      size = FONT_SIZE,
      size.unit = "pt",
      show.legend = FALSE
    ) +
    # whole proteome data (triangle)
    geom_point(
      data = wp_data,
      aes(x = protein, y = LFC_WP),
      shape = 24,
      color = "black",
      size = POINT_SIZE - 0.75,
      stroke = POINT_STROKE,
      alpha = 0.9
    ) +
    # aesthetics
    my_theme() +
    scale_fill_manual(values = regulation_colors,
                      breaks = names(regulation_colors)) +
    scale_color_manual(values = regulation_colors,
                       breaks = names(regulation_colors)) +
    scale_y_continuous(breaks = round(y_min):round(y_max),
                       limits = c(y_min, y_max)) +
    guides(fill = guide_legend(byrow = TRUE,
                               nrow = TRUE)) +
    theme(
      axis.text.x = element_text(
        color = "black",
        angle = 45,
        size = FONT_SIZE,
        family = "Arial",
        vjust = 1,
        hjust = 1
      ),
      legend.key.spacing.y = unit(-2.5, "mm"),
      legend.margin = margin(b = -15),
      legend.position = "top",
      #aspect.ratio = ratio_aspect,
      legend.title = element_text(margin = margin(b = -2)),
      legend.key = element_blank(),
      legend.title.position = "top",
      legend.key.spacing = unit(0, "mm"),
      legend.spacing.y = unit(-0, "mm"),
      legend.spacing = unit(0, "mm")
    ) +
    labs(y = substitute(
      paste('cysteine reactivity, log'[2],
            "(", n, "/", m, ")"),
      list(n = fil_condition, m = control_condition)
    ),
    x = "") +
    guides(fill = guide_legend(ncol = 2, title = paste0("Cysteine reactivity in ", fil_condition))) -> plot
  print(plot)
  save_plot(
    paste0(plot_dir, title),
    height = ht,
    width = wd
  )
  return(plt_df)
}

plot_loadings <- function(loadings_df, ggplot_obj, arrow_scaling, num_loadings = 5){
  loadings_df %>% arrange(PC1) %>% tail(num_loadings) %>% pull(variable)-> top_pc1
  loadings_df %>% arrange(PC1) %>% head(num_loadings) %>% pull(variable)-> bottom_pc1
  loadings_df %>% arrange(PC2) %>% tail(num_loadings) %>% pull(variable)-> top_pc2
  loadings_df %>% arrange(PC2) %>% head(num_loadings) %>% pull(variable)-> bottom_pc2
  
  loadings_df %>% filter(variable %in% c(top_pc1, bottom_pc1, bottom_pc2, top_pc2)) -> filtered_loadings
  loadings <- ggplot_obj + geom_segment(data= filtered_loadings, 
                      aes(x = 0, y = 0, 
                          xend = PC1 * arrow_scaling, 
                          yend = PC2 * arrow_scaling), 
                      inherit.aes = FALSE,
                    arrow = arrow(length = unit(0.1, "cm")),
                    size = LINE_WIDTH,
                    color="grey90") + 
    geom_text(data= filtered_loadings,
                    mapping = aes(label=variable,x=PC1*arrow_scaling, y=PC2*arrow_scaling), 
                    inherit.aes = FALSE, 
                    segment.size = LINE_WIDTH,
                    min.segment.length = unit(0,"mm"),
                    size=1.763, 
                    color="black",
                    #bg.color = "white",
                    segment.color = "grey") 
  return(loadings)
}


facetted_bargraph <- function(df, protein_list,shape_scale ,n_col = 5, facet_scales = "free_x", error_bars = TRUE, panel_spacing = 6){
plt <- df %>%
  filter(protein %in% protein_list) %>%
  mutate(condition = factor(x = condition, levels = names(cols)), protein = factor(x = protein, levels = protein_list)) %>%
  ggplot(aes(x = condition, y = percent_control, fill = condition)) + 
  facet_wrap(~protein, ncol = n_col, scales = facet_scales) + 
  scale_fill_manual(values = cols) + 
      geom_bar(
      stat = "summary",
      fun = "mean",
      width = 0.65,
      color = "black",
      alpha = 0.8,
      size = LINE_WIDTH,
      show.legend = FALSE
    ) +
  my_theme() + 
    geom_beeswarm(
      mapping = aes(shape = donor),
      stroke = LINE_WIDTH,
      corral.width = 1.5,
      cex = 2,
      alpha = 0.8,
      size = 1,
      #fill = "grey90",
      color = "black",
      show.legend = FALSE
    ) +
  theme(strip.background = element_blank(), strip.text = element_text(size = 6, color = "black", family = "Arial"), panel.spacing = unit(panel_spacing, "mm")) + 
      theme(
      axis.text.x = element_text(
        size = FONT_SIZE,
        color = "black",
        family = "Arial",
        angle = 45,
        vjust = 1,
        hjust = 1
      )
    ) +
    geom_hline(yintercept = 100,
               linetype = "dashed",
               size = LINE_WIDTH) +
    labs(x = NULL, y = "percent control") +
    scale_shape_manual(values = shape_scale) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))


if(error_bars){
  plt <-   plt + stat_summary(fun.data = mean_se,geom = "errorbar",width = 0.3,size = LINE_WIDTH)
}
  return(plt)
}

facetted_rc_boxplot <-
  function(proteins_to_plot,
           fig_width,
           fig_height,
            cols, fills,
           add_spacers = TRUE,
           show_legend = TRUE, df = NULL, dir = bargraph_dir) {
    df %>%
      filter(protein %in% proteins_to_plot) %>%
      mutate(
        lfc = log2(percent_control / 100),
        super_condition <-
          str_split_fixed(condition, " + ", 2)[, 1]
      ) %>%
      select(protein,
             residue,
             residue_loc,
             lfc,
             condition) -> df
    if(add_spacers){
    # add invisible spacers separating conditions D2, D8A, D8C
    for (plot_protein in proteins_to_plot) {
      map <- setNames(unique(df %>% filter(protein == plot_protein) %>% pull(residue)),
                      unique(df %>% filter(protein == plot_protein) %>% pull(residue_loc)))
      for (residue_location in names(map)) {
        for (spacer_num in c("spacer1", "spacer2")) {
          df[nrow(df) + 1, ] = list(
            "protein" = plot_protein,
            "residue" = map[residue_location],
            "residue_loc" = as.double(residue_location),
            "lfc" = 0,
            "condition" = spacer_num
          )
        }
      }
    }
    }
    
    legend_breaks = names(fills)[!grepl("spacer", names(fills))]
    df <- df %>% filter(!is.na(residue))
    df %>%
      mutate(condition = factor(df$condition, levels = names(cols))) %>%
      arrange(condition) %>%
      ggplot(aes(
        x = reorder(residue, residue_loc),
        y = lfc,
        fill = condition,
        color = condition
        #shape = condition
      )) +
      geom_boxplot(
        alpha = 0.9,
        lwd = LINE_WIDTH,
        outliers = FALSE,
        show.legend = FALSE
      ) +
      geom_jitter(
        size = POINT_SIZE - 1.5,
        alpha = 0.9,
        shape = 21,
        stroke = POINT_STROKE,
        show.legend = show_legend,
        position = position_jitterdodge(jitter.width = 0.1)
      ) +
      labs(y = substitute(paste(
        'cysteine reactivity, log'[2],
        "(FC from D2)"
      )),
      x = "") +
      guides(fill = element_blank()) +
      my_theme() +
      theme(
        strip.background = element_blank(),
        strip.text = element_text(
          color = "black",
          size = FONT_SIZE,
          family = "Arial"
        ),
        plot.title = element_text(
          size = 8,
          color = "black",
          family = "Arial",
          hjust = 0.5
        ),
        legend.title = element_blank(),
        legend.box.background = element_blank(),
        legend.position = "top",
        legend.key = element_blank(),
        legend.margin = margin(b = -14, t = -10),
        legend.key.spacing.y = unit(-3, "mm"),
        legend.text = element_text(
          size = 6,
          color = "black",
          family = "Arial",
          margin = margin(l = -4)
        ),
        legend.key.spacing.x = unit(-1, "mm")
      ) +
      scale_fill_manual(values = fills, breaks = legend_breaks) +
      scale_color_manual(values = cols, breaks = legend_breaks) +
            geom_hline(yintercept = c(1, -1),
                 linetype = "dashed",
                 size = LINE_WIDTH) +
      geom_vline(
        xintercept = 1.5:length(unique(df$residue)),
        size = LINE_WIDTH,
        linetype = "dotted"
      ) +
      facet_grid( ~ protein, space = "free") -> plt
    print(plt)
    
    save_plot(paste0(
      dir,
      paste(proteins_to_plot, collapse = "_")
    ),
    width = fig_width,
    height = fig_height)
    
    return(df)
  }

pca_plot <- function(df,
                     loadings_df,
                     x_column,
                     y_column,
                     x_lab,
                     y_lab,
                     fn,
                     arrow_scaling = 100,
                     dir = bargraph_dir,
                     use_shape = TRUE) {
  
  # Base aesthetics (always include fill and color)
  if (use_shape) {
    aes_mapping <- aes_string(
      x = x_column,
      y = y_column,
      fill = "condition",
      color = "condition",
      shape = "donor"
    )
  } else {
    aes_mapping <- aes_string(
      x = x_column,
      y = y_column,
      fill = "condition",
      color = "condition"
    )
  }
  
  df %>%
    ggplot(aes_mapping) +
    geom_point(
      size = POINT_SIZE,
      stroke = POINT_STROKE,
      alpha = 0.7,
      show.legend = TRUE,
      shape = if (!use_shape) 21 else NULL  # Set shape to 21 if use_shape is FALSE
    ) + 
    my_theme() +
    labs(x = x_lab, y = y_lab) +
    scale_fill_manual(values = cols, name = "condition") +
    scale_color_manual(values = cols, name = "condition") +
    {if (use_shape) scale_shape_manual(values = shape_scale)} +  # Only add if use_shape is TRUE
    theme(aspect.ratio = 1) +
    theme(
      legend.spacing.y = unit(0, "mm"),
      legened.key.spacing = unit(-1, "mm"),
      legend.key.height = unit(0.4, "cm"),
      legend.box.spacing = unit(0, "mm"),
      legend.margin = margin(l = 2),
      legend.title = element_text(
        size = 6,
        color = "black",
        family = "Arial",
        margin = margin()
      )
    ) -> plt
  print(plt)
  save_plot(fn, height = 2, width = 4)
  
  plot_loadings(loadings_df, plt, arrow_scaling) -> plot_with_loadings
  print(plot_with_loadings)
  save_plot(paste0(fn, "_with_loadings"),
            height = 2,
            width = 4)
}

pca_plot <- function(df,
                     loadings_df,
                     x_column,
                     y_column,
                     x_lab,
                     y_lab,
                     fn,
                     arrow_scaling = 100,
                     dir = bargraph_dir,
                     use_shape = TRUE) {
  
  # Base aesthetics (always include fill and color)
  if (use_shape) {
    aes_mapping <- aes_string(
      x = x_column,
      y = y_column,
      fill = "condition",
      color = "condition",
      shape = "donor"
    )
    geom_point_args <- list(
      size = POINT_SIZE,
      stroke = POINT_STROKE,
      alpha = 0.7,
      show.legend = TRUE
    )
  } else {
    aes_mapping <- aes_string(
      x = x_column,
      y = y_column,
      fill = "condition",
      color = "condition"
    )
    geom_point_args <- list(
      size = POINT_SIZE,
      stroke = POINT_STROKE,
      alpha = 0.7,
      show.legend = TRUE,
      shape = 21
    )
  }
  
  df %>%
    ggplot(aes_mapping) +
    do.call(geom_point, geom_point_args) +
    my_theme() +
    labs(x = x_lab, y = y_lab) +
    scale_fill_manual(values = cols, name = "condition") +
    scale_color_manual(values = cols, name = "condition") +
    {if (use_shape) scale_shape_manual(values = shape_scale)} +
    theme(aspect.ratio = 1) +
    theme(
      legend.spacing.y = unit(0, "mm"),
      legened.key.spacing = unit(-1, "mm"),
      legend.key.height = unit(0.4, "cm"),
      legend.box.spacing = unit(0, "mm"),
      legend.margin = margin(l = 2),
      legend.title = element_text(
        size = 6,
        color = "black",
        family = "Arial",
        margin = margin()
      )
    ) -> plt
  print(plt)
  save_plot(fn, height = 2, width = 4)
  
  plot_loadings(loadings_df, plt, arrow_scaling) -> plot_with_loadings
  print(plot_with_loadings)
  save_plot(paste0(fn, "_with_loadings"),
            height = 2,
            width = 4)
}

```

# Inhibitor WP PCA
```{r, fig.width = 4, fig.height=2}
cols <- c(
  "D2" = "#A1A1A1",
  "D8A + DMSO" = "#603EA6",
  "D8A + Thapsigargin" = "#8FB59D",
  "D8C + DMSO" = "#E27753",
  "D8C + BSJ" = "#4287A3",
  "D8C + ISRIB" = "#A3425B",
  "D8C + Thapsigargin" = "#3D7A54"
)
shape_scale <- c("d1" = 21,"d2" =  22,"d3" = 23,"d4" = 24)

results_dir <- '/Users/henrysanford/dev/test_data/inhibitor_treatment/both_4_donors'
#results_dir <- '/Users/henrysanford/dev/test_data/inhibitor_treatment/vardhana_2_donors'
#results_dir <- '/Users/henrysanford/dev/test_data/inhibitor_treatment/vinogradova_2_donors'

pca_dir <- paste0(results_dir, '/pca/')

df <- read.csv(paste0(pca_dir, "pca_results.csv")) %>% 
  mutate(
    donor = str_split(channel_name, "_", n = 3) %>% map_chr(2), 
    super_condition = condition %>% str_split(" ") %>% map_chr(1),
    condition = factor(condition, levels = names(cols))
    ) 


read.csv(paste0(pca_dir, "loadings_results.csv")) -> loadings_df
expl <- read_csv(paste0(pca_dir, "percent_explained.csv")) %>% pull(percent_explained)
pca_plot(
  df,
  loadings_df,
  "principal.component.1",
  "principal.component.2",
  paste0("PC1 (", expl[[1]], "%)"),
  paste0("PC2 (", expl[[2]], "%)"),
  paste0(pca_dir, "pca"),
  arrow_scaling = 125,
  use_shape = FALSE
)
df
```







```{r, fig.height=3, fig.width=6}
tp_list <- c(
    "PDCD1",
    "LAG3",
    "TIGIT",
    "HAVCR2",
    "CD39",
    "NR4A1",
    "NR4A3",
    "GZMB",
    "GSR",
    "CD62L",
    "NFATC1",
    "BHLHE40",
    "PRKCH",
    "PRKCQ",
    "SAMHD1",
    "CTLA4"
)

highglight_list <- c("PDCD1",
"LAG3",
"CTLA4",
"TIGIT",
"GZMB",
"NR4A1",
"NR4A3",
"NFATC1",
"PRKCH",
"PRKCQ")

df <- paste0(results_dir, "/bargraphs/long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)  
facetted_bargraph(df, highglight_list, shape_scale = c(21,21,21,21))

save_plot(
  fn = paste0(results_dir, "/bargraphs/selected_proteins"),
  height = 3, width = 6
)
```

```{r, fig.width=4.05, fig.height=2.15}
df <- paste0(results_dir, "/bargraphs/long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)  %>%
  filter(condition != "D8C + Gossypetin")
facetted_bargraph(df, c("TCF7", "TOX", "TBX21", "IFNG", "IL2", "TNFA"), c(21, 21, 21, 21), n_col = 5, facet_scales = "free", error_bars = TRUE)

save_plot(
  fn = paste0(results_dir, "/bargraphs/selected_proteins"),
  height = 2.15, width = 4.05
)
```

### Inhibitor bargraphs

```{r, fig.width=4.05, fig.height=2.15}
df <- paste0(results_dir, "/bargraphs/long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)  %>%
  filter(condition != "D8C + Gossypetin")
facetted_bargraph(df, c("LAG3", "PDCD1", "GZMB"), c(21, 21, 21, 21), n_col = 5, facet_scales = "free", error_bars = TRUE)

save_plot(
  fn = paste0(results_dir, "/bargraphs/selected_proteins"),
  height = 2.15, width = 4.05
)
```

```{r, fig.width=2.6, fig.height=2.15}
df <- paste0(results_dir, "/bargraphs/long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)  %>%
  filter(condition != "D8C + Gossypetin")
facetted_bargraph(df, c("GZMB", "NR4A3"), c(21, 21, 21, 21))

save_plot(
  fn = paste0(results_dir, "/bargraphs/selected_proteins"),
  height = 2.15, width = 2.6
)
```

```{r, fig.width=4.35, fig.height=2.15}
df <- paste0(results_dir, "/bargraphs/long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)  %>%
  filter(condition != "D8C + Gossypetin")
facetted_bargraph(df, c("NR4A3", "CTLA4", "NFATC1","EGR2"), c(21, 21, 21, 21), n_col = 5, facet_scales = "free", error_bars = TRUE, panel_spacing = 1)

save_plot(
  fn = paste0(results_dir, "/bargraphs/selected_proteins"),
  height = 2.15, width = 4.35
)
```


# ATF GSEA


```{r, fig.height=3, fig.width = 2}

df <- paste0(results_dir, "/gsea/gsea_results.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(comparison = paste(treatment_condition, "vs", control_condition)) %>%
  filter(Term %in% c(
    "ATF6_01_binding_motif", "ATF4_Q2_binding_motifs"
  )) %>%
  mutate(Term = case_when(
    Term == "ATF6_01_binding_motif" ~ "ATF6 binding motifs",
    Term == "ATF4_Q2_binding_motifs" ~ "ATF4 binding motifs",
    TRUE ~ Term
  )) %>%
  mutate(comparison = str_replace(comparison, " vs ", " / "))
sig_comparisons <- df %>%
  filter(`FDR q-val` < 0.05) %>%
  pull(comparison)
df <- df %>%
  #filter(comparison %in% sig_comparisons) %>%
  dplyr::select(`FDR q-val`, `NES`, Term, comparison) %>%
  arrange(NES) 
df %>%
  ggplot(aes(x = NES, y = reorder(comparison, NES), fill = -log10(`FDR q-val`))) +
  geom_col(width = 0.5, color = "black", size = LINE_WIDTH) + 
  geom_vline(xintercept = 0,size = LINE_WIDTH) +
  facet_wrap(~Term, ncol =1) + my_theme() + 
  labs(y = NULL) + theme(strip.text = element_text(size = 6)) + 
  scale_fill_continuous(low = "#B3CDE3", high = "#B10F7C", breaks = c(0.75, 1.25, 1.75)) + 
      labs(
         y = NULL,
         fill = expression("-log"[10] * "(FDR)")) +
    theme(
      axis.text.y = element_text(size = 6, family = FONT_FAMILY),
      legend.text = element_text(margin = margin(l = 1)),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(4, "mm"),
      strip.clip = "off",
      panel.grid.major = element_line(color = "grey90", size = LINE_WIDTH),
      legend.position = "bottom",
      legend.margin = margin(0, 0, -10, -10),
      strip.background = element_blank()
    ) +
  scale_x_continuous(expand = expansion(mult = c(0.2, 0.2)))

save_plot(paste0(results_dir, "/gsea/gsea_dot_plot"), height = 3, width = 2)
```






### Inhibitor bargraphs
```{r, fig.height=2, fig.width=6}

facetted_bargraph <- function(df, protein_list, shape_scale, n_col = 5){
  dodge_width <- 0.8  # Define consistent dodge width
  
  plt <- df %>%
    filter(protein %in% protein_list) %>%
    mutate(condition = factor(x = condition, levels = names(cols))) %>%
    arrange(residue_loc) %>%
    mutate(residue = factor(x = residue, levels = unique(residue))) %>%
    ggplot(aes(x = residue, y = percent_control, fill = condition)) + 
    facet_wrap(~protein, ncol = n_col, scales = "free_x") + 
    scale_fill_manual(values = cols) + 
    geom_bar(
      stat = "summary",
      fun = "mean",
      width = dodge_width,
      color = "black",
      alpha = 0.8,
      size = LINE_WIDTH,
      show.legend = FALSE,
      position = position_dodge(width = dodge_width)
    ) +
    my_theme() + 
    stat_summary(
      fun.data = mean_se,
      geom = "errorbar",
      width = 0.3,
      size = LINE_WIDTH,
      position = position_dodge(width = dodge_width)
    ) +
    geom_beeswarm(
      stroke = LINE_WIDTH,
      corral.width = 1, 
      shape =21,
      cex = 1,
      alpha = 0.8,
      size = 1,
      color = "black",
      show.legend = FALSE,
      dodge.width = dodge_width  
    ) +
    theme(
      strip.background = element_blank(), 
      strip.text = element_text(size = 6, color = "black", family = "Arial"), 
      panel.spacing = unit(6, "mm"),
      panel.border = element_blank(),
      axis.line.x = AXIS_LINE,
      axis.line.y = AXIS_LINE
    ) + 
    geom_hline(
      yintercept = 100,
      linetype = "dashed",
      size = LINE_WIDTH
    ) +
    labs(x = NULL, y = "cysteine reactivity (%D2)") +
    scale_shape_manual(values = shape_scale) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
  
  return(plt)
}

facetted_bargraph(df %>% filter(control_condition == "D2"), c("MAP2K3","MAP2K4"), c(21, 21, 21, 21))

ggsave(
  paste0(data_dir, "map2k_all_residues.svg"),
  height = 2,
  width = 6
)
```


```{r, fig.width=1.4, fig.height=2.15}

facetted_bargraph <- function(df, protein_list, shape_scale, n_col = 5){
  dodge_width <- 0.8  # Define consistent dodge width
  
  plt <- df %>%
    filter(name %in% protein_list) %>%
    mutate(condition = factor(x = condition, levels = names(cols))) %>%
    arrange(residue_loc) %>%
    mutate(residue = factor(x = residue, levels = unique(residue))) %>%
    mutate(name = factor(x = name, levels = protein_list)) %>%
    ggplot(aes(x = condition, y = percent_control, fill = condition)) + 
    facet_wrap(~name, ncol = n_col, scales = "free_x") + 
    scale_fill_manual(values = cols) + 
    geom_bar(
      stat = "summary",
      fun = "mean",
      width = 0.65,
      color = "black",
      alpha = 0.8,
      size = LINE_WIDTH,
      show.legend = FALSE,
      #position = position_dodge(width = dodge_width)
    ) +
    my_theme() + 
    stat_summary(
      fun.data = mean_se,
      geom = "errorbar",
      width = 0.3,
      size = LINE_WIDTH,
      #position = position_dodge(width = dodge_width)
    ) +
    geom_beeswarm(
      stroke = LINE_WIDTH,
      corral.width = 1, 
      shape =21,
      cex = 1,
      alpha = 0.8,
      size = 1,
      color = "black",
      show.legend = FALSE,
      #dodge.width = dodge_width  
    ) +
    theme(
      strip.background = element_blank(), 
      strip.text = element_text(size = 6, color = "black", family = "Arial"), 
      panel.spacing = unit(6, "mm"),
      #panel.border = element_blank(),
      #axis.line.x = AXIS_LINE,
      #axis.line.y = AXIS_LINE
    ) + 
          theme(
      axis.text.x = element_text(
        size = FONT_SIZE,
        color = "black",
        family = "Arial",
        angle = 45,
        vjust = 1,
        hjust = 1
      )
    ) +
    geom_hline(
      yintercept = 100,
      linetype = "dashed",
      size = LINE_WIDTH
    ) +
    labs(x = NULL, y = "cysteine reactivity (%D2)") +
    scale_shape_manual(values = shape_scale) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
  
  return(plt)
}

facetted_bargraph(df %>% filter(control_condition == "D2") %>% mutate(name = paste(protein, residue)), c("MAP2K4 C246"), c(21, 21, 21, 21))

ggsave(
  paste0(data_dir, "map2k_.svg"),
  height = 2.15,
  width = 1.4
)
```

### Non-unique peptide

```{r, fig.width=2.25, fig.height=2.15}
cysteine_name <- "MAP2K3 C207 / MAP2K6 C196"
map2k4_cysteine <- "MAP2K4 C246"
one_peptide_df <- paste0(data_dir, "KMCDFGISGYLVDSVAKT.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(name = cysteine_name, residue_loc = 207, residue = c("207")) %>%
  select(-c("SEQUENCE"))

map2k4_df <- df %>%
  filter(control_condition == "D2") %>%
  mutate(name = paste(protein, residue)) %>%
  filter(name == c(map2k4_cysteine)) %>%
  select(colnames(one_peptide_df))

rbind(map2k4_df, one_peptide_df) %>%
  facetted_bargraph(
    protein_list = c(map2k4_cysteine, cysteine_name),
    shape_scale = c(21,21,21,21)
  )

ggsave(filename = paste0(data_dir, "map2k_cystienes.svg"),width = 2.25, height = 2.15 )
```


# Mouse WP data

## PCA

```{r, fig.width=4, fig.height=2}
cols <- c(
  "Mouse D2 CD8" = "#A1A1A1",
  "Mouse D8A CD8" = "#603EA6",
  "Mouse D8C CD8" = "#E27753",
  "D8 Armstrong" = "#8b8bbc",
  "D8 Clone 13" = "#f7931e",
  "D15 Clone 13" =  "#EF3D3B",
  "TILs" = "#e8a0a7",
  "Splenic T cells" = "#a54d5b"
)

pca_dir <- '/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/pca/in_vivo/'

df <- read.csv(paste0(pca_dir, "pca_results.csv")) %>% mutate(donor = str_split(channel_name, "_", n = 3) %>% map_chr(2))

read.csv(paste0(pca_dir, "loadings_results.csv")) |> mutate(variable = toupper(variable)) -> loadings_df

expl <- read_csv(paste0(pca_dir, "percent_explained.csv")) %>% pull(percent_explained)
pca_plot(
  df,
  loadings_df,
  "principal.component.1",
  "principal.component.2",
  paste0("PC1 (", expl[[1]], "%)"),
  paste0("PC2 (", expl[[2]], "%)"),
  paste0(pca_dir, "pca"),
  arrow_scaling = 200,
  use_shape = FALSE
)

```



## Bargraphs

```{r, fig.width=4.5, fig.height=3}
bargraph_dir <- '/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/bargraphs/'

df <- paste0(bargraph_dir, "long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)
protein_list <- c("PDCD1","LAG3","CTLA4","TIGIT","GZMB","NR4A1","NR4A3","NFATC1","PRKCH","PRKCQ")

mouse_protein_list <-  c("Pdcd1","Lag3","Ctla4","Tigit","Gzmb","Nr4a1","Nr4a3","Nfatc1","Prkch","Prkcq")
facetted_bargraph(df, mouse_protein_list, shape_scale = c(21), n_col = 3)

save_plot(
  fn = paste0(bargraph_dir, "selected_proteins"),
  height = 3, width = 4.5
)
```


```{r, fig.width=4.2, fig.height=2.15}

mouse_protein_list <-  c("Pdcd1","Lag3","Havcr2")
facetted_bargraph(df, mouse_protein_list, shape_scale = c(21), n_col = 3, error_bars = FALSE,  facet_scales = "free", panel_spacing = 2)

save_plot(
  fn = paste0(bargraph_dir, "selected_proteins"),
  height = 2.15, width = 4.2
)

```


## Volcano plots

```{r, fig.width=2, fig.height=4}

group_hierarchy_redox <- function(df){
  df %>%
    mutate(group = 
           case_when(
             nucleotide_metabolism ~ "Nucelotide metabolism",
             redox_related ~ "Redox related",
             cell_cycle ~ "Cell cycle",
             TRUE ~ "Other"
           )) %>% 
    return()
}
group_hierarchy_etc <- function(df){
  df %>%
    mutate(group = 
           case_when(
             respiratory_complex_v ~ "Respiratory_complex_V",
             `respiratory_complex_i-iv` ~ "Respiratory_complex_I_IV",
             mitochondrial ~ "Mitochondrial",
             TRUE ~ "Other"
           )) %>% 
    mutate(group = factor(group,levels = rev(names(color_code)))) %>%
    arrange(group) %>%
    return()
}
group_hierarchy_peroxisome <- function(df){
  df %>%
    mutate(group = 
           case_when(
             peroxisome ~ "Peroxisome",
             TRUE ~ "Other"
           )) %>% 
    mutate(group = factor(group,levels = rev(names(color_code)))) %>%
    arrange(group) %>%
    return()
}

# read volcano data
df <- "/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/long_volcano_df.csv" %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(comparison = paste0(condition, " vs. ",control_condition))
all_comparisons <- unique(df$comparison)


# redox volcano plot
df <- group_hierarchy_redox(df) %>% mutate(log10_pval = `-log10_pval`)
color_code <- c(
  "Nucelotide metabolism" = "#bade28",
  "Redox related" = "#440154",
  "Cell cycle" = "#21918c",
  "Other" = 'gray88'
)
volcano_dir <- "/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/"
row = c(group = "Multiple groups")
run_all_comparisons(
  df, row, all_comparisons, color_code, c("Gpx1","Gpx4","Gsr","Sod2","Selenow", "Cdk4","Cdk6","Cdkn2a","Cdc25c","Chek2","Entpd1","Cd38","Ak4","Nt5dc2")
)


color_code <- c(
  "Respiratory_complex_V" = "#bade28",
  "Respiratory_complex_I_IV" = "#440154",
  "Mitochondrial" = "#21918c",
  "Other" = 'gray88'
)
df <- group_hierarchy_etc(df) %>% mutate(log10_pval = `-log10_pval`)
row = c(group = "Complex_I-V")


run_all_comparisons(
  df, row, all_comparisons, color_code, c("a")
)

color_code <- c(
  "Peroxisome" = "#21918c",
  "Other" = 'gray88'
)



df <- group_hierarchy_peroxisome(df) %>% mutate(log10_pval = `-log10_pval`)
row = c(group = "Peroxisome")
run_all_comparisons(
  df, row, all_comparisons, color_code, c("a")
)
```

### GSEA of PC loadings

```{r, fig.width=3.5, fig.height=4}
pca_dir <- "/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/pca/"
clean_term <- function(term) {
  str_split_fixed(term, "_", 2)[2] -> term
  #gsub("\\b([a-z])", "\\U\\1", tolower(term), perl=TRUE) -> term
  str_replace_all(term, "_", " ") -> term
  str_wrap(term, 40) -> term
  return(term)
}

replot_pca <- function(pca_dir){
 for (pc in c("PC1", "PC2")) {
  paste0(pca_dir, pc,
         "/gseapy.gene_set.prerank.report.csv") %>%
    read_csv(show_col_types = FALSE) %>%
    mutate(
      Term = lapply(Term, FUN = clean_term) %>% unlist(),
      `-log10(FDR)` = -log10(`FDR q-val`)
    ) %>%
    arrange(abs(NES)) -> df
  
  max_score  <- max(df$`-log10(FDR)`[is.finite(df$`-log10(FDR)`)], na.rm = TRUE)
  df$`-log10(FDR)`[is.infinite(df$`-log10(FDR)`)] <- 5.5
  
  # color palette
  down_df <- df %>%
    filter(`FDR q-val` < 0.01) %>%
    filter(NES  < 0) |> 
    tail(10)
  up_df <- df |> 
    filter(`FDR q-val` < 0.01) %>%
    filter(NES  > 0) |> 
    tail(10) 
  down_df |>  rbind(up_df) |> 
    ggplot(aes(
      x = NES,
      y = reorder(Term, NES),
      fill = `-log10(FDR)`
    )) +
    geom_col(color = "black",
             size = LINE_WIDTH,
             width = 0.5) +
    my_theme() +
    geom_vline(xintercept = 0,
               color = "black",
               size = LINE_WIDTH) +
    labs(x = paste0(pc, " NES"),
         y = NULL,
         fill = expression("-log"[10] * "(FDR)")) +
    theme(
      axis.text.y = element_text(size = 5, family = FONT_FAMILY),
      legend.text = element_text(margin = margin(l = 1)),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(4, "mm"),
      #panel.border = element_blank(),
      panel.grid.major = element_line(color = "grey90", size = LINE_WIDTH),
    ) + 
      #scale_fill_distiller(palette = "BuPu", direction = 1) -> plt
    scale_fill_continuous(low = "#B3CDE3", high = "#B10F7C", limits = c(1.5, 5.5)) -> plt
  print(plt)
  save_plot(
    paste0(pca_dir, pc, "/top_enriched_sets"),
    width = 3.5,
    height = 4
  )
} 
}

replot_pca(pca_dir)

```

```{r}
# in vivo only
pca_dir <- "/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/pca/in_vivo/"
replot_pca(pca_dir)

```



# Human vs Mouse

### Dot plot

```{r, fig.height=4.5, fig.width=2}
data_dir <- "/Users/henrysanford/dev/test_data/texh_revisions/mouse_wp/output/"
proteins_of_interest <- c("BATF","CTLA4","CXCL13","EGR2","ENTPD1","GZMB","HAVCR2","IL7R","IRF4","LAG3","NFATC1","NR4A1","TBX21","TCF7","TNFRSF9","BHLHE40", "IFNG", "IL2", "TNF", "TIGIT", "TOX", "PDCD1", "CD38", "HLA-DRA","HLA-DRB1","HLA-DRB3", "CD244", "NT5E", "NR4A3", "NFKB2","LTB","BATF","CTLA4","TNFRSF9","EGR2","CD200","TNFRSF4")
df <- data_dir |> 
paste0("combined_long_volcano_data.csv") |> 
  read_csv(show_col_types = FALSE) |> 
  filter(
    # only show comparisons to D2
    control_condition %in% c("D2", "Mouse D2 CD8"),
    # proteins to plot
    human %in% proteins_of_interest
    ) |> 
  mutate(dataset = case_when(
    dataset == "human" ~ "human\nin vitro",
    condition %in% c("Mouse D8A CD8","Mouse D8C CD8" ) ~ "mouse\nin vitro",
    TRUE ~ "mouse\nin vivo"
  ))   |> group_by(human) |>
  #filter(n_distinct(dataset) == n_distinct(df$dataset)) |>
  ungroup() |> 
  mutate(condition = factor(condition, levels = c("D8A", "D8C", "Mouse D8A CD8", "Mouse D8C CD8", "D8 Armstrong", "D8 Clone 13", "D15 Clone 13", "TILs"))) |> 
    # temporary fix because the orthology database is too broad
  filter(!(`mouse, laboratory` %in% c("Gzma","Gzmc","Gzmd","Gzme","Gzmf","Gzmg" )))
  
protein_order <- df |> filter(condition == "D8C") |> arrange(log2_FC) |> pull(human) |> unique()

df |> 
  mutate(human = factor(human, levels = c("PDCD1",protein_order))) |> 
  ggplot(aes(x = condition, y = human, size = -log10(p_value), fill = log2_FC)) +
  geom_point(shape =21, stroke = LINE_WIDTH) + 
  facet_grid(cols = vars(dataset), scales = "free_x", space = "free") + 
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", limits = c(-3, 3), oob = scales::squish) + 
  my_theme() + 
      theme(
      strip.background = element_blank(), 
      strip.text = element_text(size = 6, color = "black", family = "Arial"), 
      panel.spacing = unit(0, "mm"),
      panel.grid.major = element_line(color = "grey90", size = LINE_WIDTH)
    ) + 
          theme(
      axis.text.x = element_text(
        size = FONT_SIZE,
        color = "black",
        family = "Arial",
        angle = 45,
        vjust = 1,
        hjust = 1
      ),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.box = "vertical"
    )  +
  labs(x = NULL, y = NULL, size = expression("-log"[10]*"(p-value)"), fill = expression("-log"[2]*"(FC from mouse or human D2)")) +
      guides(
      fill = guide_colorbar(
        barwidth = 4,
        barheight = 0.6,
        title.position = "top",
        title.hjust = 0.5,
        order =1
      ),
      size = guide_legend(
        direction = "horizontal",
        title.position = "top",
        title.hjust = 0.5,
        nrow = 1,
        order =2
      )
    ) +
  scale_size_continuous(range = c(0.5, 4)) 
df |> mutate(human = factor(human, levels = unique(protein_order)))
ggsave(
  paste0(data_dir, "mouse_vs_human_dot_plot.svg"),
  height = 4.5, 
  width = 2
)
```


```{r, fig.width=4, fig.height=2}

input_dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Exhaustion manuscript/01_RNA-sequencing/13524/'
paste0(
  input_dir, "gsea_enrichment_curve/chronic_activation_score/input/GMDF.gmx"
) %>% 
  read_csv(show_col_types = FALSE,skip = 1) %>% 
  pull("> Genes unregulated in exhausted T-cells from Aviv Regev") -> exhaustion_markers_43


human_control <- "D2"
human_condition <- "D8C"
mouse_control <- "Mouse D2 CD8"
mouse_condition <- "Mouse D8C CD8"
df <- data_dir |> 
paste0("combined_long_volcano_data.csv") |> 
  read_csv(show_col_types = FALSE) |> 
  filter(
    (control_condition ==human_control & condition == human_condition) | (control_condition ==mouse_control & condition == mouse_condition),
    #human %in% proteins_of_interest
    ) |> 
  drop_na(human) |> 
  select(human, condition, control_condition, log2_FC,  dataset, peroxisome, cell_cycle) |> 
  select(-c(condition, control_condition)) |> 
  rename(protein = human) |> 
  pivot_wider(
    id_cols = c(protein, peroxisome, cell_cycle), 
    names_from = dataset,
    values_from = log2_FC,
      values_fn = first
  ) |> drop_na() |> 
  mutate(exhausiton_marker = protein %in% exhaustion_markers_43) |> 
  mutate(exhausiton_marker= factor(exhausiton_marker, levels = c(TRUE, FALSE)))
lbl_df <- df |> filter(protein %in% proteins_of_interest)

df |> 
  arrange(desc(exhausiton_marker)) |> 
  ggplot(
    aes(x = human, y = mouse, fill = exhausiton_marker, label = protein)
  ) + 
  geom_point(shape = 21, alpha = 0.5, stroke = LINE_WIDTH, size = 1) + 
  my_theme() +
  theme(aspect.ratio = 1) + 
  geom_text_repel(data = lbl_df, size = FONT_SIZE_MM, min.segment.length = 0, nudge_y = 1, segment.size = LINE_WIDTH) + 
  scale_fill_manual(values = c("#21918c", "grey90"))


exhaustion_markers_43
```


# Human TIL data

```{r}
cols <- c(
'Human D0 CD3' = "lightgrey",
 'Human D2 CD3' = "#A1A1A1",
 'Human D8A CD3' = "#603EA6",
 'Human D8C CD3' = "#E27753",
 'TIL 1' = "#ac6468",
 'TIL 2' = "#a74245",
 'TIL 3' = "#d39e9c",
'TIL 4' = "#7a2e31", 
 'TIL 5' = "#e8c4c3"
)
```

### PCA

```{r, fig.width=4, fig.height=2}

data_dir <- "/Users/henrysanford/dev/test_data/texh_revisions/new_human_til/output/"
pca_dir <- paste0(data_dir, 'pca/')

df <- read.csv(paste0(pca_dir, "pca_results.csv")) %>% mutate(donor = str_split(channel_name, "_", n = 3) %>% map_chr(2))

read.csv(paste0(pca_dir, "loadings_results.csv")) -> loadings_df

expl <- read_csv(paste0(pca_dir, "percent_explained.csv")) %>% pull(percent_explained)
pca_plot(
  df,
  loadings_df,
  "principal.component.1",
  "principal.component.2",
  paste0("PC1 (", expl[[1]], "%)"),
  paste0("PC2 (", expl[[2]], "%)"),
  paste0(pca_dir, "pca"),
  arrow_scaling = 150,
  use_shape = FALSE
)


```

```{r}
replot_pca("/Users/henrysanford/dev/test_data/texh_revisions/new_human_til/output/pca/")
```

```{r}
replot_pca("/Users/henrysanford/dev/test_data/texh_revisions/new_human_til/output/pca/in_vivo/")
```



```{r, fig.width = 4, fig.height=2}

plot_loadings <- function(loadings_df, ggplot_obj, arrow_scaling, num_loadings = 5){
  loadings_df %>% arrange(PC1) %>% tail(num_loadings) %>% pull(variable)-> top_pc1
  loadings_df %>% arrange(PC1) %>% head(num_loadings) %>% pull(variable)-> bottom_pc1
  loadings_df %>% arrange(PC3) %>% tail(num_loadings) %>% pull(variable)-> top_pc2
  loadings_df %>% arrange(PC3) %>% head(num_loadings) %>% pull(variable)-> bottom_pc2
  
  loadings_df %>% filter(variable %in% c(top_pc1, bottom_pc1, bottom_pc2, top_pc2)) -> filtered_loadings
  loadings <- ggplot_obj + geom_segment(data= filtered_loadings, 
                      aes(x = 0, y = 0, 
                          xend = PC1 * arrow_scaling, 
                          yend = PC3 * arrow_scaling), 
                      inherit.aes = FALSE,
                    arrow = arrow(length = unit(0.1, "cm")),
                    size = LINE_WIDTH,
                    color="grey90") + 
    geom_text(data= filtered_loadings,
                    mapping = aes(label=variable,x=PC1*arrow_scaling, y=PC3*arrow_scaling), 
                    inherit.aes = FALSE, 
                    segment.size = LINE_WIDTH,
                    min.segment.length = unit(0,"mm"),
                    size=1.763, 
                    color="black",
                    #bg.color = "white",
                    segment.color = "grey") 
  return(loadings)
}

pca_plot <- function(df,
                     loadings_df,
                     x_column,
                     y_column,
                     x_lab,
                     y_lab,
                     fn,
                     arrow_scaling = 100,
                     dir = bargraph_dir,
                     use_shape = TRUE) {
  
  # Base aesthetics (always include fill and color)
  if (use_shape) {
    aes_mapping <- aes_string(
      x = x_column,
      y = y_column,
      fill = "condition",
      color = "condition",
      shape = "donor"
    )
  } else {
    aes_mapping <- aes_string(
      x = x_column,
      y = y_column,
      fill = "condition",
      color = "condition"
    )
  }
  
  df %>%
    ggplot(aes_mapping) +
    geom_point(
      size = POINT_SIZE,
      stroke = POINT_STROKE,
      alpha = 0.7,
      show.legend = TRUE,
      shape = if (!use_shape) 21 else NULL  # Set shape to 21 if use_shape is FALSE
    ) + 
    my_theme() +
    labs(x = x_lab, y = y_lab) +
    scale_fill_manual(values = fills, name = "condition") +
    scale_color_manual(values = cols, name = "condition") +
    {if (use_shape) scale_shape_manual(values = shape_scale)} +  # Only add if use_shape is TRUE
    theme(aspect.ratio = 1) +
    theme(
      legend.spacing.y = unit(0, "mm"),
      legened.key.spacing = unit(-1, "mm"),
      legend.key.height = unit(0.4, "cm"),
      legend.box.spacing = unit(0, "mm"),
      legend.margin = margin(l = 2),
      legend.title = element_text(
        size = 6,
        color = "black",
        family = "Arial",
        margin = margin()
      )
    ) -> plt
  print(plt)
  save_plot(fn, height = 2, width = 4)
  
  plot_loadings(loadings_df, plt, arrow_scaling) -> plot_with_loadings
  print(plot_with_loadings)
  save_plot(paste0(fn, "_with_loadings"),
            height = 2,
            width = 4)
}

cols <- c(
'Human D0 CD3' = "black",
 'Human D2 CD3' = "#A1A1A1",
 'Human D8A CD3' = "#603EA6",
 'Human D8C CD3' = "#E27753",
 'TIL 1' = "#ac6468",
 'TIL 2' = "#a74245",
 'TIL 3' = "#d39e9c",
'TIL 4' = "#7a2e31", 
 'TIL 5' = "#e8c4c3"
)

fills <- c(
'Human D0 CD3' = "transparent",
 'Human D2 CD3' = "#A1A1A1",
 'Human D8A CD3' = "#603EA6",
 'Human D8C CD3' = "#E27753",
 'TIL 1' = "#ac6468",
 'TIL 2' = "#a74245",
 'TIL 3' = "#d39e9c",
'TIL 4' = "#7a2e31", 
 'TIL 5' = "#e8c4c3"
)


pca_plot(
  df |> mutate(condition = factor(condition, levels  = names(cols))),
  loadings_df,
  "principal.component.1",
  "principal.component.3",
  paste0("PC1 (", expl[[1]], "%)"),
  paste0("PC3 (", expl[[3]], "%)"),
  paste0(pca_dir, "pca"),
  arrow_scaling = 175,
  use_shape = FALSE
)
```



### Bargraphs

```{r, fig.width=4.2, fig.height=2.15}
bargraph_dir <- paste0(data_dir, 'bargraphs/')

df <- paste0(bargraph_dir, "long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)
protein_list <- unique(c("PDCD1","LAG3","CTLA4","TIGIT","GZMB","NR4A1","NR4A3","NFATC1","ENTPD1","HAVCR2","BHLHE40","CTLA4","TNFRSF9","EGR2","CXCL13","TIGIT"," TOX","TCF7"))

facetted_bargraph(df, protein_list, n_col = 3, shape_scale = c(21, 22))

save_plot(
  fn = paste0(bargraph_dir, "selected_proteins"),
  height = 3.5, width = 4
)
```

```{r, fig.width=3.2, fig.height=2.15}


bargraph_dir <- paste0(data_dir, 'bargraphs/')

df <- paste0(bargraph_dir, "long_percent_control.csv") %>%
  read_csv(show_col_types = FALSE)
protein_list <- unique(c("LAG3","GZMB"))

facetted_bargraph(df, protein_list, n_col = 3, shape_scale = c(21, 21), facet_scales = "free")

save_plot(
  fn = paste0(bargraph_dir, "lag3_gzmb"),
  height = 2.15, width = 3.2
)
```

```{r, fig.width=1.2, fig.height=2.15}
protein_list <- unique(c("PDCD1"))

facetted_bargraph(df, protein_list, n_col = 3, shape_scale = c(21, 22), facet_scales = "free", error_bars = FALSE, )

save_plot(
  fn = paste0(bargraph_dir, "pdcd1"),
  height = 2.15, width = 1.2
)

```


### Volcano plot

```{r}

nudge_x = 0.7
make_condition_plot_multi <- function(row, 
                                condition, 
                                volcano_data, 
                                control_condition, 
                                color_code,
                                highlight_proteins = c(), 
                                highlight_color = '#440154'){
  x_var <- paste0("log2_FC")
  y_var <- paste0("log10_pval")
  change_var = paste0("Regulation")
  name_var = "protein"
  group <- row["group"]
  
  change_var = paste0("Regulation")
  y_max <- max(abs(volcano_data$`-log10_pval`)) * 1.2
  x_max <- max(abs(volcano_data$log2_FC)) * 1.5
  nudge_x <- 2
  # build label dataframe
  labels <- list()
  y <- y_max * (15.5/16)
  x <- x_max * (4.7/5.3)
  print(names(color_code))
  for (direction in c("Significant Up", "Significant Down")){
   for (bio_group in names(color_code)){
      volcano_data[volcano_data[[change_var]] == direction &
                           volcano_data[["group"]] == bio_group,] %>% 
       nrow() -> protein_count
     r <- c("direction" = direction,
            "bio_group" = bio_group, 
            "count" = protein_count, 
            "x" = x,
            "y" = y)
     labels[[paste0(bio_group, direction)]] <- r
     y <- y - (y_max * (1.5/16))
   }
    y <- y_max * (15.5/16)
    x <- x * -1
  }
  
  do.call("rbind", labels) %>% as.data.frame() -> count_label_df
  title <- paste0(control_condition," vs ", condition)
  
  alphas_binary <- c(0.4,0.9)
  
  sig <- volcano_data[ volcano_data[[change_var]] %in% c("Significant Up", "Significant Down") , ]
  
  volcano_data %>% mutate(significant = uniprot %in% sig$uniprot) -> volcano_data
  sig$significant <- TRUE
  label_data <- sig[ sig[[name_var]] %in% highlight_proteins , ]
  set.seed(12345)
  volcano_data %>%
    arrange(group) %>%
    ggplot(aes_string(x=x_var,
               y=y_var,
                fill = "group",
               alpha = "significant")) +
    geom_point(data = volcano_data %>% filter(group == "Other"), 
               size = 1,
               shape=21,
               stroke=POINT_STROKE / 2,
               col="black") +
    geom_point(data = volcano_data %>% filter(group != "Other"),
              size = 1,
               shape=21,
               stroke=POINT_STROKE / 2,
               col="black") +
    geom_point(data = label_data,
              size = 1,
               shape=21,
               stroke=POINT_STROKE / 2,
               col="black") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", size=LINE_WIDTH) +
    geom_vline(xintercept = c(-FC_CUTOFF, FC_CUTOFF), linetype = "dashed", size=LINE_WIDTH) +
    geom_text_repel(
      data = label_data[label_data[[change_var]] == "Significant Down",],
      aes_string(x = x_var,
      y = y_var,
      color = "group",
      # label with gene name instead of uniprot
      label = name_var),
      direction    = "both",
      #color = "black",
      box.padding = 1,
      nudge_x = -nudge_x,
      xlim = c(NA, -1),
      size =FONT_SIZE_MM,
      #bg.color = "white",
      min.segment.length = 0,
      segment.size = LINE_WIDTH,
      inherit.aes = FALSE,
      show.legend = FALSE) +
    geom_text_repel(
      data = label_data[label_data[[change_var]] == "Significant Up",],
      aes_string(x = x_var,
      y = y_var,
      color = "group",
      # label with gene name instead of uniprot
      label = name_var),
      direction    = "both",
      #color = "black",
      box.padding = 1,
      nudge_x = nudge_x,
      xlim = c(1,NA),
      size = FONT_SIZE_MM,
      #bg.color = "white",
      min.segment.length = 0,
      segment.size = LINE_WIDTH,
      inherit.aes = FALSE,
      show.legend = FALSE) +
  geom_label(
    data = count_label_df,
    mapping = aes(x = as.numeric(x), 
                  y = as.numeric(y), 
                  label = count, 
                  fill = bio_group),
    color = "white", 
    size=6,
    size.unit="pt",
    show.legend = FALSE,
    inherit.aes = FALSE) +
    theme_test() +
    labs(
      x = substitute(paste('protein expression, log'[2], "(", n, '/', m, ")"), list(n = condition, m = control_condition)),
      y = substitute(paste('-log'[10], '(p-value)'))
    ) +
    scale_fill_manual(values=color_code) +
        scale_color_manual(values=color_code) +
    scale_alpha_manual(values = alphas_binary, guide="none") +
    my_theme() +
  theme(strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.title=element_blank(),
      legend.box.background = element_blank(),
      legend.position = "top",
      legend.justification.top = "top",
      legend.justification = "center",
      legend.box.just = "center",
      aspect.ratio=1,
      legend.key = element_blank(),
      legend.margin = margin(l = -5, b = -5.5),
      legend.box.margin = margin(l = -5, b = -0.8),
      legend.key.spacing.y = unit(-0.9, "cm"),
      legend.key.spacing.x = unit(-0.05, "cm")) + 
    guides(fill = guide_legend(ncol = 2, 
                               )) +
    ylim(0,y_max) +
    xlim(-x_max,x_max) -> v_plot
  print(v_plot)
  
  save_plot(paste0(volcano_dir, "functional_groups_highlighted/", group, "/", title), width=2, height=4)
}

run_all_comparisons <- function(df, row, all_comparisons, color_code, highlight_proteins){
  for (comp in all_comparisons){
  print(comp)
  condition <- comp %>% str_split(" vs. ") %>% map_chr(1) 
  control <- comp %>% str_split(" vs. ") %>% map_chr(2) 
  
  make_condition_plot_multi(
    row, 
    condition, 
    df %>% filter(comparison == comp),
    control, 
    color_code = color_code,
    highlight_proteins = c("ayo"), 
    highlight_color = "#440154")
}
}

volcano_dir <- paste0(data_dir, "volcano_plots/")
# read volcano data
df <- paste0(data_dir, "long_volcano_df.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(comparison = paste0(condition, " vs. ",control_condition))

max_finite <- max(df$log2_FC[is.finite(df$log2_FC)])
min_finite <- min(df$log2_FC[is.finite(df$log2_FC)])

df$log2_FC[df$log2_FC == Inf]  <- max_finite
df$log2_FC[df$log2_FC == -Inf] <- min_finite

all_comparisons <- unique(df$comparison)


# redox volcano plot
df <- group_hierarchy_redox(df) %>% mutate(log10_pval = `-log10_pval`)
color_code <- c(
  "Nucelotide metabolism" = "#bade28",
  "Redox related" = "#440154",
  "Cell cycle" = "#21918c",
  "Other" = 'gray88'
)
row = c(group = "Multiple groups")
run_all_comparisons(
  df, row, all_comparisons, color_code, c("GPX1","GPX4","GSR","SOD2","SELENOW", "CDK4","CDK6","CDKN2A","CDC25C","CHEK2","ENTPD1","CD38","AK4","NT5DC2")
)

# etc volcano plot
color_code <- c(
  "Respiratory_complex_V" = "#bade28",
  "Respiratory_complex_I_IV" = "#440154",
  "Mitochondrial" = "#21918c",
  "Other" = 'gray88'
)
df <- group_hierarchy_etc(df) %>% mutate(log10_pval = `-log10_pval`)
row = c(group = "Complex_I-V")


run_all_comparisons(
  df, row, all_comparisons, color_code, c("a")
)

color_code <- c(
  "Peroxisome" = "#21918c",
  "Other" = 'gray88'
)



df <- group_hierarchy_peroxisome(df) %>% mutate(log10_pval = `-log10_pval`)
row = c(group = "Peroxisome")
run_all_comparisons(
  df, row, all_comparisons, color_code, c("a")
)



```

```{r}
df |> filter(log2_FC == max_finite) |> pull(protein) |> unique() |> paste(collapse = " ")
```


# T-cell expression plots

bar graph showing expression changes for CSNK2A1, CSNK2A2, and CSNK2B (D2, D8A, D8C);
bar graph showing expression changes for MAF1 (D2, D8A, D8C);
Reactivity change for BRF2_C370 in acute and chronic conditions;

```{r, fig.width=2.3, fig.height=2}
wp_percent_control_df <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Exhaustion manuscript/02_Mass-spectrometry data/01_whole_proteome/02_wp_analysis/20240821_6rep_2MC_two_replicate_var_filter/03_combined_files/20240821_6rep_2MC_two_replicate_var_filter/02_combfiles_percctrl_wp.csv'

results_dir <- '/Users/henrysanford/Dropbox @RU Dropbox/Vinogradova Laboratory/Exhaustion manuscript/02_Mass-spectrometry data/01_whole_proteome/02_wp_analysis/20240821_6rep_2MC_two_replicate_var_filter/04_results/addl_bargraphs/'

cols <-c(
  "D2" = "#A1A1A1", 
  "D4A" = "#ABA2D6", 
  "D8A" = "#603EA6", 
  "D4C" = "#FBB31B", 
  "D8C" = "#E27753"
)

preprocess_wp_file <- function(filepath){
  df <- filepath |> 
  read_csv(show_col_types = FALSE) |> 
  select(-contains("pepNum")) |> 
  select(-c("...1")) |> 
  pivot_longer(cols = contains("processed"), values_to = "percent_control") |> 
  separate_wider_delim(
    col = name,
    delim = "_",
    names = c("condition","donor", "technical_replicate"),
    too_many = "drop"
  )
  
  return(df)
}

original_wp_bargraph <- function(list_proteins){
# format wp data for plotting
plt <-  wp_percent_control_df|> 
  preprocess_wp_file() |> 
  filter(condition %in% c("D2","D8A","D8C")) |> 
  facetted_bargraph(protein_list = list_proteins, shape_scale = rep(21, 6)) + 
  labs(x = NULL, y = "protein expression (%D2)") + 
  theme(axis.text.x = element_text(size = FONT_SIZE, family = FONT_FAMILY, color = "black", angle = 0, hjust = 0.5, vjust = NULL), panel.spacing = unit(1, "mm"), panel.border = element_blank(), axis.line.x = AXIS_LINE, axis.line.y = AXIS_LINE)
  print(plt)
  return(plt)
}
original_wp_bargraph(c("CSNK2A1", "CSNK2A2", "CSNK2B"))

ggsave(paste0(results_dir, "CNSK_bargraph.svg"), height = 2, width = 2.3)

```

```{r, fig.height=2, fig.width=1.1}
original_wp_bargraph(c("MAF1"))
ggsave(paste0(results_dir,"MAF1_bargraph.svg"), height = 2, width = 1.1)

```
